### 2025-03-01 code/manuscript.Rmd
push for second draft review

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy, or equipment failures occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../paper"
    )
  })
---

```{r setup, include = FALSE, cache = FALSE}

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      # cache = T,
                      dpi = 300,
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/manuscript/figs/")

# str(knitr::opts_chunk$get()) # see for all options

```

```{r prep, include = FALSE, cache = FALSE}
#### LIBRARIES ####
require(pacman)

# load packages using pacman
pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, gridExtra, patchwork, RColorBrewer)

# turn off scientific notation
options(scipen = 999, digits = 15)

# set directory
here::i_am("manuscript.rmd")

# set default theme for ggplot
theme_set(theme_minimal())

# define base ggplot theme
theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
             plot.background = element_rect(fill = "white", colour = NA),
             panel.grid.minor = element_blank(),
             panel.grid.major = element_blank(),
             axis.text = element_text(size = 10),
             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
             strip.background = element_blank())

# colors
ls_colors <- c("Baseline" = "#c6dbef",
               "Measured baseline" = "#c6dbef",
               "Adjusted baseline" = "#c6dbef",
               "Projected baseline" = "#2171b5",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#66c2a4",
               "Randomized\n(50/50)" = "#66c2a4",
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Randomized\n(20/80)" = "#ccece6",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

# parameters
ctr_params <- list(peak_hours = 10:16,
                   chwl_perc = 0.25,
                   step_perc = 0.08,
                   conv_swt = 6,
                   weather_knots = c(15, 25),
                   swt_knots = c(12, 7),
                   coe_peak = 0.8,
                   coe_off = 1.2,
                   enable_temp = 8)

occ_params <- list(change_start = 5,
                   change_end = 8,
                   change = 20)

#### FUNCTIONS ####
# Function defined to read downloaded tmy files
get_tmy <- function(all_sites, readfile_path){

  df_tmy <- data.frame()

  for (site in all_sites){
    df <- read_csv(paste0(readfile_path, "tmy/", str_glue("{site}.epw")),
                   skip = 8, col_types = "ddddd-d---------------------------------",
                   col_names = c("year", "month", "day", "hour", "min", "tmy")) %>%
      mutate(year = 2017,
             time = ymd_h(paste(paste(year, month, day, sep = "-"), hour, sep = " ")),
             temp = tmy) %>%
      dplyr::select(time, temp) %>% 
      mutate(site = site)

    df_tmy <- rbind(df_tmy, df)
  }

  return(df_tmy)
}

# Function defined to add chwst reset intervention
run_reset <- function(df_baseline){

  mean <- mean(df_baseline$base_eload, na.rm = T) * ctr_params$chwl_perc

  grad <- (ctr_params$swt_knots[2] - ctr_params$swt_knots[1]) / 
    (ctr_params$weather_knots[2] - ctr_params$weather_knots[1])

  interc <- ctr_params$swt_knots[2] - (ctr_params$weather_knots[2] * grad)

  df_interv <- df_baseline %>% 
    mutate(swt = t_out * grad + interc, 
           chwl = mean,
           hour = hour(datetime)) %>% 
    mutate(swt = ifelse(swt > ctr_params$swt_knots[1], ctr_params$swt_knots[1], ifelse(swt < ctr_params$swt_knots[2], ctr_params$swt_knots[2], swt)), 
           temp_savings = ifelse(t_out >= ctr_params$enable_temp, (swt - ctr_params$conv_swt) * ctr_params$step_perc, 0), 
           time_adj = ifelse(hour %in% ctr_params$peak_hours, ctr_params$coe_peak, ctr_params$coe_off), 
           perc_savings = temp_savings * time_adj, 
           savings = chwl * perc_savings, 
           interv_eload = base_eload - savings) %>% 
    select(datetime, base_eload, interv_eload, t_out)

  return(df_interv)
}

# Function defined to interpolate NAs
run_interpo <- function(df_all){

  na_counts <- df_all %>%
    mutate(date = date(timestamp)) %>%
    group_by(date) %>%
    summarize(na_hours = sum(is.na(eload)))

  # Filter out days with more than half of the hours having NAs
  valid_days <- na_counts %>%
    filter(na_hours <= 12) %>%
    pull(date)

  df_filtered <- df_all %>%
    filter(date(timestamp) %in% valid_days)

  df_filtered <- df_filtered %>%
    mutate(across(c(eload, t_out), ~ zoo::na.approx(., na.rm = FALSE))) %>% 
    rename(datetime = timestamp, 
           base_eload = eload)

  return(df_filtered)
}

# Function defined to adjust the plot scale
get_scale <- function(eload, range = 2){

  min_y <- mean(eload, na.rm = T) - range * sd(eload, na.rm = T)
  max_y <- mean(eload, na.rm = T) + range * sd(eload, na.rm = T)

  return(c(min_y, max_y))
}
```

```{r readdata, include=FALSE}

#### READ DATA ####
df_loc <- read.csv("../readfiles/loc_map.csv")

# stable dataset
readfile_stable <- str_glue("../readfiles/stable/")
fig_path = "../figs/manuscript/"

df_energy_stable <- read_rds(paste0(readfile_stable, "df_energy.rds"))
df_weather_stable <- read_rds(paste0(readfile_stable, "df_weather.rds"))
df_sprt_all_stable <- read_rds(paste0(readfile_stable, "df_sprt_all.rds"))
df_seq_fs_stable <- read_rds(paste0(readfile_stable, "df_seq_fs.rds"))
df_fs_stable <- read_rds(paste0(readfile_stable, "df_fs.rds"))
df_cont_stable <- read_rds(paste0(readfile_stable, "df_cont.rds"))
df_fs_null_stable <- read_rds(paste0(readfile_stable, "df_fs_null.rds"))
df_seq_fs_null_stable <- read_rds(paste0(readfile_stable, "df_seq_fs_null.rds"))
df_model_acc_stable <- read_rds(paste0(readfile_stable, "df_model_acc.rds"))
df_fs_tmy_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy.rds"))
df_fs_tmy_null_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy_null.rds"))
df_interval_drop_stable <- read_rds(paste0(readfile_stable, "df_interval_drop.rds"))
df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))

all_sites_stable <- df_energy_stable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_stable <- df_energy_stable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_stable <- df_energy_stable %>%
  select(name) %>%
  distinct(name)

# variable dataset
readfile_variable <- str_glue("../readfiles/variable/")

df_energy_variable <- read_rds(paste0(readfile_variable, "df_energy.rds"))
df_weather_variable <- read_rds(paste0(readfile_variable, "df_weather.rds"))
df_sprt_all_variable <- read_rds(paste0(readfile_variable, "df_sprt_all.rds"))
df_seq_fs_variable <- read_rds(paste0(readfile_variable, "df_seq_fs.rds"))
df_fs_variable <- read_rds(paste0(readfile_variable, "df_fs.rds"))
df_cont_variable <- read_rds(paste0(readfile_variable, "df_cont.rds"))
df_fs_null_variable <- read_rds(paste0(readfile_variable, "df_fs_null.rds"))
df_seq_fs_null_variable <- read_rds(paste0(readfile_variable, "df_seq_fs_null.rds"))
df_model_acc_variable <- read_rds(paste0(readfile_variable, "df_model_acc.rds"))
df_fs_tmy_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy.rds"))
df_fs_tmy_null_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy_null.rds"))
df_interval_drop_variable <- read_rds(paste0(readfile_variable, "df_interval_drop.rds"))
df_interval_keep_variable <- read_rds(paste0(readfile_variable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_keep.rds"))

all_sites_variable <- df_energy_variable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_variable <- df_energy_variable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_variable <- df_energy_variable %>%
  select(name) %>%
  distinct(name)

# read functions
function_path <- "../functions/"
source(paste0(function_path, "model_fit.R"))
source(paste0(function_path, "model_pred.R"))
source(paste0(function_path, "prepost_plot.R"))

S_building <- nrow(df_fs_tmy_stable)
V_building <- nrow(df_fs_tmy_variable)
A_building <- S_building + V_building
```

# Introduction

## Conventional M&V

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the 
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. This limitation
reduces the feasibility of rapid M&V and complicates the quantification
of estimation uncertainty, thus impacting the accuracy and timeliness of
savings assessments.

## Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we propose a novel M&V method
that adopts the randomized crossover design, a gold standard from
medical and agricultural studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. Another improvement is that we
proposed a sequential evaluation framework and defined stopping criteria
to end the M&V if the target effect is detected. This is to avoid
unnecessary measurement collection over the full 2-year M&V cycle. The
full framework is detailed in a previous published study with all
stopping criteria and example usecases outlined [@raftery2024]. In
summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially right after test begins.
In addition, once the desired savings target is achieved, analysts can
terminate the M&V and switch to 100% intervention. The key advantage of
randomization is that if control strategies are sampled with equal
probability, the influence of other confounding factors such as
occupancy change (or commonly known as non-routine events) is likely to
be evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention's impact.

## BDG2 dataset

The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly use the metered electricity and outdoor weather measurements for
running the M&V methods.

## Literature review

### Whole building approach

ASHRAE Guidelinge 14 and IPMVP provides options for whole building M&V
use cases. For code compliance using prescriptive option, 12 months of
baseline and 12 months of postretrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline model fitting accuracy should have CV(RMSE) lower than 20%.
Thus, most research related to M&V for whole-building approach focuses
on the accuracy of baseline modeling, exploring model performance from
simple regression models to more complex machine learning techniques.
One study reviewed various models suitable for M&V applications as well
as selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that baseline model might deteriorate (i.e.
becomes 'stale') over an extended period of pre- and post-analysis.

### Non-routine events impact

Non-routine events during M&V commonly refers to unexpected changes in a
building that influences its energy usage. These changes can greatly
affect measured energy consumption in buildings and are typically
unrelated to the intervention strategy. Therefore, those changes are
considered 'static factors' respect to measured independent variables. A
common non-routine event in energy-saving M&V projects is a change in
occupancy or a significant shift in occupant behavior, equipment run
time, and operating conditions (e.g. set points, lighting and
ventilation levels). However, current standards or guidelines only
provides general approach for consideration. For example, IPMVP requires
facility owner and the M&V analyst periodically perform inspections of
all equipment and operations during reporting period, which is labor
intensive and error-prone. ASHRAE guideline 14 recommends performing
engineering calculations or computer software simulations to adjust
post-retrofit baseline. Additional, it is very rare to have access to
all measurements such as occupancy, and thus the analyst normally assume
those factors remain unchanged throughout the study. Specifically,
studies investigating the effects of demand response on building energy
efficiency commonly uses linear interpolation to estimate
counter-factual baseline [@beil_round-trip_2015;
@keskar_commercial_2020], and one study points out that it is inaccurate
to assume no change in the operating conditions during the test period
[@huang_experimental_2023]. One likely encountered non-routine event due
to operation condition change is filter clogging in air handling units
due to particle accumulation. This can cause supply fans to gradually
consume more energy to maintain required duct static pressure
[@feng_newly_2019; @zhai_full-scale_2017]. If M&V analysts are unaware
of such changes and lack an appropriate adjustment method (e.g.,
replacing filters before the intervention begins), the savings could be
underestimated as increased energy use is incorrectly attributed to the
intervention rather than the mechanical issue. Some studies realized the
limitation of current M&V methods, which only consider adjusting for
outdoor weather, is insufficient and emphasized the importance of
requiring matched comparison groups to control for exogenous factors
beyond weather differences when comparing between baseline and
intervention [@demand_side_analytics_population_2022;
@huang_experimental_2023]. However, those methods are still unable to
accurately quantify uncertainties due to changing baseline.

## Objectives

As mentioned, the goal of an M&V project is to determine the
effect---typically energy savings---of an switchable energy-efficient
intervention. The goal is to determine, for a large sample of buildings,
how much more accurately the novel M&V method would estimate the savings
of such an intervention compared to conventional M&V, and how much more
quickly it would reach a result. An example of such an intervention is a
control retrofit developed by a software-as-a-service company that
adjusts the chilled water plant's supply water temperature, which can be
commonly found in the literature [@duarte_field_2023;
@lee_simulationoptimization_2012; @qiu_chilled_2022; @jin2007;
@taylor_optimizing_2012]. In our case, we make it even simpler by
adjusting setpoint based on outdoor weather conditions with more
description shown in Section \@ref(apply-virtue-intervention).
Therefore, we defined the M&V scenario as follows:

"A company aims to sell its supply temperature reset control software
package to a customer, in this case, the building owner, with a
promotion that it will reduce the building's electricity usage. If the
building owner decides to purchase the service, the company agrees to
charge a service fee based on a percentage of the measured energy
savings."

As required by the M&V scenario, we assessed the performance of both the
conventional and the novel randomized M&V methods by estimating the
intervention energy savings for all valid buildings in the BG2 dataset.
By conducting such analysis, we aim to:

1.  Compare the energy saving estimation accuracy between the
    conventional and the randomized method. This study extends the
    comparison to a large sample of buildings, covering a variety of
    types and climate zones. The comparison metrics include both
    estimation accuracy and M&V completion timeline.

2.  Verify the enhanced robustness of the randomized method. By using
    realistic measurements from real-world buildings, which include
    various sources of noise, we aim to reflect the challenges faced by
    building analysts in real projects. As we will perform numercial
    simulation on the existing dataset, the ground-truth savings can be
    calculated as the reference for comparison.

3.  Open-source implementation of the proposed randomized M&V method
    using a public dataset. We ensured the reproducibility of the method
    by making the analysis code open source including randomized
    schedule generation, sequential statistical analysis, energy
    modeling and normalized saving calculation. In addition, we also
    included codes for extended use cases of the randomized method, such
    as changing sampling ratio and sampling intervals. Using the
    available open resources, building analysts should be able to
    seamlessly integrate and apply them in their own M&V projects.

# Method

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section.

```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

knitr::include_graphics(paste0(fig_path, "flowchart.png"))
```

## Filter and clean dataset

In this study, we extracted the electricity measurements from the BDG2
dataset and filter out all qualified buildings based on the following
criteria:

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to 1.5 months of missing days.

2.  Mean electricity usage \> 0 kWh: target buildings should have active
    electricity usage.

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease of building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5%
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrenceberkeleynationallab; @mathew_big-data_2015].

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water set point reset
    control.

```{r stablesite, fig.cap = "Site summary of the stable building subset (counts < 10 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_variable$type)

stable_set <- df_energy_stable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
   left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "location")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(location, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of building numbers and types")
```

For all qualified buildings, we further added one more stringent
criterion to explore all buildings with very stable electricity usage
between the two years:

1.  No statistical significant difference (P-value \> 0.05) between the
    two-year electricity usage, which are sometimes assumed to be the
    case for whole-building measurement and verification.

As a results, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption made
by conventional M&V is overly simplified. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events and according to the statistics shown here, is
commonly observed in real applications.

## Apply virtue intervention

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller is activated when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

```

We mapped the chilled water supply temperature reset to the electrical
energy savings as: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
further consumes around 50% of the HVAC electricity. While this
assumption largely simplifies the diverse energy usage across various
building types, for the scope of this paper, we assume that 25% of the
total building electricity is used by the chilled water plant, $E_{CHW}$
[@us2012commercial]. Typically, the savings from an intervention are not
proportional to the building's hourly electricity usage, which is
generally the challenge for M&V. To address this, we mapped the
resulting electricity savings as a percentage of the plant's normal
operation, calculated as its mean electricity usage over the two-year
period. This percentage is influenced by factors such as outdoor
temperature ($OAT$), intervention supply water temperature
($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
of the day ($\delta_{occ}$, binary indicator whether during peak hours
from 9 AM to 4 PM).

$$
\text{perc_savings = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building and were applied uniformly in the filtered dataset.
Interested readers can also change the parameters in our open-source
code to simulate different scenarios.

|                        |                                                                  |       |
|--------------|--------------------------------------------|--------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Table 1. Parameters for calculating the intervention savings.

## Run M&V methods

We described more in detail the workflow of both conventional and
randomized M&V methods in the previous study [@raftery2024]. The
conventional method is a simple pre- and post- comparison on a 12-month
baseline and 12-month intervention timeline. The randomized method
defines sampling requirements as:

-   Use a daily sampling interval with the sampling time at midnight
    each day.

-   Block by day of the week with a block period of 12 weeks.

and stopping criteria as:

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but stopping
    criteria enables an early stop at the end of satisfied blocking
    period.

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data sampled by both strategies.

-   Test for no carryover effect using a t-test with a p-value not
    exceeding a defined significance threshold of 0.05.

-   90% confidence that energy savings exceed or do not exceed 0% using
    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
    and calculated SPRT statistics either falls below the lower
    threshold or exceeds the upper threshold.

-   test with an equal sampling ratio (50% baseline, 50% intervention).

# Results

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step represents 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both at the stopping criteria and over a two-year period).", fig.height=12, fig.width=12}

#### GA ####
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p_accuracy <- ggarrange(p_top, p_middle, 
                        ncol = 1, nrow = 2,
                        labels = c("a)", "b)"),
                        common.legend = T, 
                        legend = "bottom")

# Timeline plot
df_time <- df_sprt_all_variable %>% 
  filter(seq != "final") %>% 
  bind_rows(df_sprt_all_stable %>% filter(seq != "final")) %>% 
  select(name, seq, n_weeks)

count <- list()
n <- 1
for (i in seq(0, 36, by = 3)){

  df_sprt <- df_time %>% 
    filter(seq == "sprt",
           n_weeks <= i)

  df_eob <- df_time %>% 
    filter(seq == "eob",
           n_weeks <= i)

  df_temp <- df_time %>% 
    filter(seq == "temp", 
           n_weeks <= i)

  count[[n]] <- tibble("n_weeks" = i, 
                       "eob" = nrow(df_eob), 
                       "temp" = nrow(df_temp), 
                       "sprt" = nrow(df_sprt))

  n <- n + 1
}

count <- bind_rows(count)

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 30, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall comparison of conventional and randomized M&V")
```

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison. Subplots a) and b) calculate the savings estimation error as
the absolute deviation from the true savings (i.e. $|Acc|$) and shows
the results distribution using conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot a) shows the savings estimated from measured weather
condition and subplot b) shows the savings normalized on typical
meteorological weather file after using TOWT model. As shown clearly in
the plots, the conventional M&V method exhibits an average deviation of
6% in savings estimation, whereas the proposed randomized method
demonstrates significantly smaller deviations. If the analyst stops
immediately after satisfying all stopping criteria, the deviation is
reduced to approximately 2%. Extending the M&V period to match the
conventional method further improves accuracy to 1%. Furthermore, by
comparing subplot (a) and (b), we observe a minor improvement (less than
1%) in estimation accuracy due to model fitting. This is because the
weather condition over the two years for all selected locations are
similar. Thus the effect of adjusting temperature as an independent
variable in the model is not significant. Overall the results suggest
that projecting the baseline into the post-retrofit period that only
considers weather variation is inefficient. Subplot (c) presents the
overall timeline comparison of the two M&V methods. The results indicate
that a significant portion achieve accurate M&V results within 24 weeks
and all buildings satisfy the stopping criteria within 36 weeks, mostly
influenced by weather variability. Meanwhile, the red arrow highlights
the excess time required by the conventional M&V method, which only
leads to less reliable results.

We also plotted the mean error in Figure \@ref(fig:meanerr) and we
noticed that the uncertainty range associated with the conventional
method is significantly larger compared to the randomized method that
only takes 36 weeks according to subplot c) in the Figure
\@ref(fig:abstract).

```{r meanerr, fig.cap="Distribution of mean deviation from the ground-truth savings calculated by the two M&V methods", fig.height=6, fig.width=12}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Mean error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Mean error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_middle, 
          ncol = 1, nrow = 2,
          labels = c("a)", "b)")) +
  plot_annotation(title = "Mean error in savings estimation by different M&V methods")

percentiles <- df_TW_A %>%
  filter(method == "conv_diff") %>% 
  summarise(
    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
  )
```

Additionally in the literature we found another M&V method study using
over 500+ commercial buildings, but their study scope was to compare a
variety of M&V baseline models from simple week mean to more complex
machine learning methods using the normalized mean bias error
(NMBE)[@granderson_accuracy_2016]. Since we also calculated the
normalized savings estimation error, we compared our normalized accuracy
of 600+ buildings with TOWT model and TMY weather with their TOWT model
prediction assessment in the following table. As each percentile is
closely matched, our results are verified.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-----------------|--------------------------------------|-----------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

: Table 1. Savings accuracy comparison with a similar literature

## Non-routine events impact

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect", fig.height=6, fig.width=12}

#### MEAN-NULL ####
rand_eob_S <- df_seq_fs_null_stable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_null_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_null_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_null_variable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_null_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_null_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                                "conv_diff" = "Conventional", 
                                "rand_eob_diff" = "Randomized", 
                                "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 6), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Mean effect in fractional savings",
       title = "Mean effect of non-routine events when using different M&V methods", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-18, 18)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
the selected buildings are affected by non-routine events, which is the
reason why we observe baseline electricity usage changes throughout the
two year. To evaluate the reliability of the two M&V methods in blocking
the effect of non-routine events, we repeated the M&V run on the
original dataset prior to the application of the reset control. In other
words, as there is no intervention applied or the intervention is known
to have no impact on whole-building electricity usage, a more reliable
M&V method should detect closer to 0 kW savings using the measurements.
Therefore, the $Acc$ here are calculated as:

$$
Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$

Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
time of satisfying all stopping criteria and at the end of two-year
period) produced consistent results. We included the absolute deviation
($|ACC|$) in the appendix. Similarly, stopping early can lead to
slightly larger uncertainty similar to Figure \@ref(fig:abstract).
However, as for conventional method, the uncertainty range is much
larger than the randomized method and it shows on average the deviation
is 1.3% meaning a building analyst can 'detect' 1.3% energy savings when
there isn't any using the conventional M&V method.

# Discussion

## TOWT modeling accuracy

To evaluate modeling accuracy, we used the Coefficient of Variation of
Root-Mean Squared Error (CV(RMSE)) as the primary error metric. Since
CV(RMSE) is a normalized measure, it enables direct comparison across
different model fitting results. According to ASHRAE Guideline 14,
whole-building baseline models should achieve a CV(RMSE) below 25% for
computing savings or 30% when less than 12 months' of measurements are
available [@ashrae_ashrae_2023]. In our study, Figure \@ref(fig:towtacc)
compares the model fitting accuracy of the two M&V methods and it shows
around 87.5% of the models fitted in the randomized M&V method are code
compliance while 80% for the conventional M&V method. As the quantity of
the training set is the same, this highlights the impact of data
sampling on model fitting accuracy, and for TOWT, the distribution of
outdoor weather conditions sampled in the training dataset is
influential. The conventional method trains the model using the entire
first-year dataset, if the location is heating or cooling dominant, the
data-driven model is trained using colder or warmer outdoor conditions
leading to biased model coefficients. Although the random sampling in
our study does not specifically block the outdoor temperature factor,
the CV(RMSE) distribution shown in the figure indicates a significant
improvement in model fitting accuracy.

```{r towtacc, fig.cap="TOWT model fitting accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}

cvrmse <- df_model_acc_stable %>% 
  bind_rows(df_model_acc_variable) %>% 
  pivot_longer(c(conv, rand), 
               names_to = "method", 
               values_to = "cvrmse") %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                              "rand" = "Randomized", 
                              "conv" = "Conventional"))

result <- t.test(filter(cvrmse, method == "Randomized")$cvrmse, 
                 filter(cvrmse, method == "Conventional")$cvrmse, 
                 exact = TRUE,
                 alternative = "less",
                 mu = 0,
                 conf.int = TRUE, 
                 conf.level = 0.90,
                 var.equal = TRUE)$p.value

cvrmse %>% 
  ggplot(aes(x = method, y = cvrmse, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  annotate(geom = "text", 
           x = 1.5, 
           y = 40, 
           size = 4, 
           label = paste0("P value = ", round(result, digits = 3))) +
  geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
  geom_text(data = . %>% 
              group_by(method) %>% 
              summarise(mean = mean(cvrmse, na.rm = T)) %>% 
              ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
  geom_hline(yintercept = 25, color = "red", lty = "dashed") +
  annotate(geom = "text", 
           x = 0.5, 
           y = 27, 
           size = 4, 
           color = "red", 
           label = paste0(25, " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %")) +
  coord_cartesian(ylim = c(0, 50)) +
  labs(x = NULL, 
       y = NULL, 
       color = NULL, 
       title = "TOWT model fitting accuracy", 
       subtitle = "CV(RMSE)") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

Additionally, we also noticed in our study that despite the regression
models capture well on the mean energy consumption of the building but
the tend to underestimate the 15-min daily peak load [@granderson2021],
which is not useful for assessing savings for demand response events.

## Sampling interval impact

In certain situations, particularly those involving large thermal mass
in active thermal energy storage systems, there is a known time-lagged
effect associated with the intervention. This means that the effect of
one strategy may persist after switching to another. To mitigate such
carryover effects, building analysts may choose to increase the sampling
interval and drop non-consecutive days in the switchback experiment. For
example, with a hot water tank that is charged by an intervention
strategy, residual heat may still be present when the next strategy
begins, influencing subsequent measurements, particularly the
performance measured on the first day (i.e. a non-consecutive day). In
such cases, increasing the sampling interval (e.g., sample strategies
every 2 or 3 days instead of daily) and discard measurements from those
non-consecutive days can eliminate potential carryover effect. However,
this approach reduces the total data collected: for instance, sampling
every 3 days would drop 1/6 of measurements and sampling every 2 day
means would remove 1/4 of measurements. Therefore, by increasing the
sampling interval, there is a reduction in statistical power due to less
randomness, and when decreasing the sampling interval, there is a
accuracy penalty from dropping more non-consecutive days. To quantify
such trade-off, \@ref(fig:absinterval) shows the accuracy of different
sampling intervals and consequence of dropping the non-consecutive days
in the dataset over the entire two-year period.

```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}

#### ABS-INT ####
# dropped version
df_interval_keep_stable %<>% 
  left_join(df_fs_stable %>% filter(method == "rand"), by = c("name", "site")) %>% 
  select(-c("scenario", "method")) %>% 
  rename("interval_fs_1" = "savings") %>% 
  left_join(df_fs_tmy_stable, by = c("name", "site")) %>% 
  select(-conv) %>% 
  rename("interval_tmy_1" = "rand")

df_interval_keep_variable %<>% 
  left_join(df_fs_variable %>% filter(method == "rand"), by = c("name", "site")) %>% 
  select(-c("scenario", "method")) %>% 
  rename("interval_fs_1" = "savings") %>% 
  left_join(df_fs_tmy_variable, by = c("name", "site")) %>% 
  select(-conv) %>% 
  rename("interval_tmy_1" = "rand")

df_drop <- df_interval_drop_stable %>% 
  bind_rows(df_interval_drop_variable) %>% 
  select(name, contains("fs")) %>% 
  rename("drop_1" = "interval_fs_1", 
         "drop_2" = "interval_fs_2", 
         "drop_3" = "interval_fs_3", 
         "drop_7" = "interval_fs_7")

df_keep <- df_interval_keep_stable %>% 
  bind_rows(df_interval_keep_variable) %>% 
  select(name, contains("fs")) %>% 
  rename("keep_1" = "interval_fs_1",
         "keep_2" = "interval_fs_2", 
         "keep_3" = "interval_fs_3", 
         "keep_7" = "interval_fs_7")

df_MW_A <- df_drop %>% 
  left_join(df_keep, by = c("name")) %>%
  left_join(rbind(df_fs_stable %>% 
              filter(scenario == "ref" & method == "true"), 
              df_fs_variable %>% 
              filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(drop_1 = abs(drop_1 - savings),
         drop_2 = abs(drop_2 - savings), 
         drop_3 = abs(drop_3 - savings), 
         drop_7 = abs(drop_7 - savings), 
         keep_1 = abs(keep_1 - savings),
         keep_2 = abs(keep_2 - savings), 
         keep_3 = abs(keep_3 - savings), 
         keep_7 = abs(keep_7 - savings)) %>% 
  pivot_longer(cols = contains("drop") | contains("keep"),  
               names_to = c("group", "interval"),
               names_pattern = "(drop|keep)_(\\d+)",
               values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_jitter(alpha = 0.8, 
              size = 0.5, 
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(group, interval) %>%
              summarise(mean = mean(diff, na.rm = T)) %>% 
              ungroup(), 
            aes(x = interval, y = mean, group = group,
                label = paste0(round(mean, digits = 1), " %")), 
            position = position_dodge(width = 0.75), 
            color = "white") +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 10)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY version
df_drop <- df_interval_drop_stable %>% 
  bind_rows(df_interval_drop_variable) %>% 
  select(name, contains("tmy")) %>% 
  rename("drop_1" = "interval_tmy_1",
         "drop_2" = "interval_tmy_2", 
         "drop_3" = "interval_tmy_3", 
         "drop_7" = "interval_tmy_7")

df_keep <- df_interval_keep_stable %>% 
  bind_rows(df_interval_keep_variable) %>% 
  select(name, contains("tmy")) %>% 
  rename("keep_1" = "interval_tmy_1",
         "keep_2" = "interval_tmy_2", 
         "keep_3" = "interval_tmy_3", 
         "keep_7" = "interval_tmy_7")

df_TW_A <- df_drop %>% 
  left_join(df_keep, by = c("name")) %>%
  left_join(rbind(df_fs_stable %>% 
              filter(scenario == "ref" & method == "true"), 
              df_fs_variable %>% 
              filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(drop_1 = abs(savings - drop_1),
         drop_2 = abs(savings - drop_2), 
         drop_3 = abs(savings - drop_3), 
         drop_7 = abs(savings - drop_7), 
         keep_1 = abs(savings - keep_1),
         keep_2 = abs(savings - keep_2), 
         keep_3 = abs(savings - keep_3), 
         keep_7 = abs(savings - keep_7)) %>% 
  pivot_longer(cols = contains("drop") | contains("keep"),  
               names_to = c("group", "interval"),
               names_pattern = "(drop|keep)_(\\d+)",
               values_to = "diff")

p_bottom <- df_TW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_jitter(alpha = 0.8, 
              size = 0.5, 
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(group, interval) %>% 
              summarise(mean = mean(diff, na.rm = T)) %>% 
              ungroup(), 
            aes(x = interval, y = mean, group = group,
                label = paste0(round(mean, digits = 1), " %")), 
            position = position_dodge(width = 0.75), 
            color = "white") +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TMY weather")) +
  coord_cartesian(ylim = c(0, 10)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_bottom, 
          ncol = 1, nrow = 2, 
          align = "hv", 
          common.legend = T, 
          legend = "bottom") +
  plot_annotation(title = "Absolute error in savings estimation of ranodmized M&V by different sampling intervals")

```

To quantify the impact of reduced randomness, all measurements are kept
when calculating the distribution shown in the darker colored set. The
results suggests a gradual increase in error from 1% to 1.8% when
sampling from daily to weekly intervals. To quantify the combined impact
of data loss due to preventing carryover effect, we generated the
lighter colored set, which drops all non-consecutive days. As a result,
we noticed a significant increase in estimaion error when sampling
daily, which lead to a similar accuracy compared to weekly intervals.
Therefore, considering this trade-off, we recommend using a three-day
sampling interval if the carryover effect is likely but expected to last
less than one day (as more sampling is required for a two-day
swithback), and then normalizing the estimation via energy modeling. We
also included the mean deviation distribution and also the accuracy
distribution calculated after all stopping criteria satisfied in the
supplementary material and the results shows the same indication.

## Sampling ratio impact

Another advantage of using randomized M&V is the flexibility of changing
sampling ratio after the target savings detected. For instance, the
building owner can continue sampling at a 50%/50% ratio between the
baseline and the intervention to further reduce the uncertainty
associated with the savings. Alternatively, they could switch to 100%
intervention to maximize cost savings, though this approach risks the
baseline becoming outdated, which is another trade-off to consider. A
middle-ground approach would be to sample at an 20%/80% ratio or 10%/90%
between the baseline and the intervention. To compare those possible
choices to a building owner, figure \@ref(fig:abscont) shows the results
of changing the original 50%/50% sampling ratio after satisfying all
stopping criteria and continuing a new ratio till the end of the
two-year period. We also provided the reference case when the building
owner continues with no change. The results shows as the sampled
baseline days and intervention days become unbalanced, the savings
estimation error grows from 1% to 5%. This is reasonable since sampling
at 10%/90% is similar to the conventional M&V method. We included the
mean deviation version in the supplementary material.

```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy", fig.height=6, fig.width=12}

#### ABS-CONT ####
rand_cont_S <- df_cont_stable %>% 
  select(-cont_tmy) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs)) %>% 
  select(name, diff, ratio)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         ratio = 0) %>% 
  select(name, diff, ratio)

df_MW_S <- rbind(rand_cont_S, rand_final_S)

# plot for the variable subset
rand_cont_V <- df_cont_variable %>% 
  select(-cont_tmy) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs)) %>% 
  select(name, diff, ratio)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         ratio = 0) %>% 
  select(name, diff, ratio)

df_MW_V <- rbind(rand_cont_V, rand_final_V)

# plot for combined dataset
rand_cont_A <- bind_rows(rand_cont_V, rand_cont_S) %>% 
  select(name, rand_cont_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rbind(df_MW_S, df_MW_V)

df_MW_A %>% 
  mutate(ratio = as.factor(ratio), 
         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)")) %>% 
  ggplot(aes(x = ratio, y = diff, fill = "Randomized")) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(ratio) %>% 
              summarise(mean = mean(diff)) %>% 
              ungroup(), 
            aes(x = ratio, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       title = "Absolute error in savings estimation by different sampling ratio",
       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Limitations

We identify two key limitations in this study:

1.  The simulated intervention remains somewhat generic and simplistic
    considering the diverse building types and climate zones in the BDG2
    dataset. As the primary focus is to accurately detect the
    intervention (or any intervention) effect so for simplicity, we
    applied the parameters listed in Table 1 uniformly across all
    buildings. Yet, in some cases, raising the water temperature by 1°C
    might yield more or less than 8% in electricity savings due to
    target buildings' demand flexibility. For example, one field
    validation of HVAC optimization shows additional savings might be
    limited if other best practice controls are already implemented
    [@granderson2018].
2.  For the main analysis, we assumed no carryover effect and a daily
    sampling interval would suffice for most commercial buildings, but
    exceptions exist. Based on the measurement from the dataset, we have
    limited knowledge of how builings thermal mass respond to the
    proposed chilled water reset strategy. However, in the discussion,
    we addressed such concern by comparing different sampling strategies
    and the impact of dropping measurements from non-consecutive days.

# Conclusion

This research demonstrated the application of a novel whole-building
measurement and verification (M&V) method, comparing its performance to
the conventional approach outlined in ASHRAE Guideline 14 using a large,
open-source commercial building dataset. The proposed M&V method
leverages the randomized experimental design concept from other
scientific fields, along with statistical sequential inference
techniques, to determine when target savings are detected. We used a
virtual control retrofit case---resetting the chilled water setpoint
based on outdoor weather conditions---and applied it to over 600
filtered commercial buildings. By comparing the savings estimations of
the conventional method with the novel randomized method, we found that
the randomized approach provides faster and more robust savings
estimations.

Specifically, we showed that throughout 11 different locations assessed
in this study, the randomized M&V can provide a saving estimation by 36
weeks (with the majority finishes by 24 weeks) once all stopping
criteria satisfied. In contrast, the conventional method requires a full
range of baseline and intervention measurements under normal operating
conditions, typically taking 6-9 months for each phase. Most
importantly, Furthermore, we verified that with reduced M&V timeline the
randomized method can estimate savings more accurately by showing that
the absolute error is only 1 - 2 % while for the two-year conventional
method, the estimation error could be 6% in a typical building. We also
evaluated the impact of non-routine events on the proposed M&V method
and the results show that baseline changes in the post retrofit period
can deviate savings estimated using the conventional method. As a
comparison, we found those events have very negligible impact on the
savings estimated using the randomized method, demonstrating ideal
reliability for realistic use cases.

We also discussed the impact of model fitting accuracy when using the
two methods and we found a significant improvement in our method due to
random sampling. We also provide detailed assessment on using different
sampling interval and sampling ratio, those results provide useful
considerations to building analysts when generating randomized schedules
for a variety of use cases.

# CRediT authorship contribution statement

**Aoyu Zou**: Conceptualization, Data curation, Formal analysis,
Methodology, Investigation, Software, Writing - original draft,
Writing - review & editing. **Paul Raftery**: Conceptualization, Formal
analysis, Methodology, Investigation, Project administration,
Supervision, Writing - review & editing. **Stefano Schiavon**:
Conceptualization, Methodology, Investigation, Supervision, Writing -
review & editing. **Carlos Duarte**: Methodology, Investigation,
Supervision, Writing - review & editing. **Gail Brager**: Supervision,
Writing - review & editing.

# Reproducibility

Reproducible example with analysis code is available (MIT license) at
<https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.

# Declaration of competing interest

The authors declare that they have no known competing financial
interests or personal relationships that could have appeared to
influence the work reported in this paper.

# Acknowledgements

# References