### 2025-03-22 code/manuscript.Rmd
Uploaded revised manuscript, added sampling interval timeline

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events including static changes and gradual changes denoted as operational drift occurs during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability. We also assessed the impact of dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../paper"
    )
  })
---

```{r setup, include = FALSE, cache = FALSE}

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      # cache = T,
                      dpi = 300,
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/manuscript/figs/")

# str(knitr::opts_chunk$get()) # see for all options

```

```{r prep, include = FALSE, cache = FALSE}
#### LIBRARIES ####
require(pacman)

# load packages using pacman
pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, gridExtra, patchwork, RColorBrewer)

# turn off scientific notation
options(scipen = 999, digits = 15)

# set directory
here::i_am("manuscript.rmd")

# set default theme for ggplot
theme_set(theme_minimal())

# define base ggplot theme
theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
             plot.background = element_rect(fill = "white", colour = NA),
             panel.grid.minor = element_blank(),
             panel.grid.major = element_blank(),
             axis.text = element_text(size = 10),
             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
             strip.background = element_blank())

# colors
ls_colors <- c("Baseline" = "grey10",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#6baed6",
               "Randomized\n(50/50)" = "#6baed6",
               "Randomized\n(20/80)" = "#6baed6", 
               "Randomized\n(10/90)" = "#6baed6", 
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Dropped (24 months)" = "#4292c6",
               "Kept (24 months)" = "#08519c",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

# parameters
ctr_params <- list(peak_hours = 10:16,
                   chwl_perc = 0.25,
                   step_perc = 0.08,
                   conv_swt = 6,
                   weather_knots = c(15, 25),
                   swt_knots = c(12, 7),
                   coe_peak = 0.8,
                   coe_off = 1.2,
                   enable_temp = 8)

occ_params <- list(change_start = 5,
                   change_end = 8,
                   change = 20)

#### FUNCTIONS ####
# Function defined to read downloaded tmy files
get_tmy <- function(all_sites, readfile_path){

  df_tmy <- data.frame()

  for (site in all_sites){
    df <- read_csv(paste0(readfile_path, "tmy/", str_glue("{site}.epw")),
                   skip = 8, col_types = "ddddd-d---------------------------------",
                   col_names = c("year", "month", "day", "hour", "min", "tmy")) %>%
      mutate(year = 2017,
             time = ymd_h(paste(paste(year, month, day, sep = "-"), hour, sep = " ")),
             temp = tmy) %>%
      dplyr::select(time, temp) %>% 
      mutate(site = site)

    df_tmy <- rbind(df_tmy, df)
  }

  return(df_tmy)
}

# Function defined to add chwst reset intervention
run_reset <- function(df_baseline){

  mean <- mean(df_baseline$base_eload, na.rm = T) * ctr_params$chwl_perc

  grad <- (ctr_params$swt_knots[2] - ctr_params$swt_knots[1]) / 
    (ctr_params$weather_knots[2] - ctr_params$weather_knots[1])

  interc <- ctr_params$swt_knots[2] - (ctr_params$weather_knots[2] * grad)

  df_interv <- df_baseline %>% 
    mutate(swt = t_out * grad + interc, 
           chwl = mean,
           hour = hour(datetime)) %>% 
    mutate(swt = ifelse(swt > ctr_params$swt_knots[1], ctr_params$swt_knots[1], ifelse(swt < ctr_params$swt_knots[2], ctr_params$swt_knots[2], swt)), 
           temp_savings = ifelse(t_out >= ctr_params$enable_temp, (swt - ctr_params$conv_swt) * ctr_params$step_perc, 0), 
           time_adj = ifelse(hour %in% ctr_params$peak_hours, ctr_params$coe_peak, ctr_params$coe_off), 
           perc_savings = temp_savings * time_adj, 
           savings = chwl * perc_savings, 
           interv_eload = base_eload - savings) %>% 
    select(datetime, base_eload, interv_eload, t_out)

  return(df_interv)
}

# Function defined to interpolate NAs
run_interpo <- function(df_all){

  na_counts <- df_all %>%
    mutate(date = date(timestamp)) %>%
    group_by(date) %>%
    summarize(na_hours = sum(is.na(eload)))

  # Filter out days with more than half of the hours having NAs
  valid_days <- na_counts %>%
    filter(na_hours <= 12) %>%
    pull(date)

  df_filtered <- df_all %>%
    filter(date(timestamp) %in% valid_days)

  df_filtered <- df_filtered %>%
    mutate(across(c(eload, t_out), ~ zoo::na.approx(., na.rm = FALSE))) %>% 
    rename(datetime = timestamp, 
           base_eload = eload)

  return(df_filtered)
}

# Function defined to adjust the plot scale
get_scale <- function(eload, range = 2){

  min_y <- mean(eload, na.rm = T) - range * sd(eload, na.rm = T)
  max_y <- mean(eload, na.rm = T) + range * sd(eload, na.rm = T)

  return(c(min_y, max_y))
}
```

```{r readdata, include=FALSE}

#### READ DATA ####
df_loc <- read.csv("../readfiles/loc_map.csv")

# stable dataset
readfile_stable <- str_glue("../readfiles/stable/")
fig_path = "../figs/manuscript/"

df_energy_stable <- read_rds(paste0(readfile_stable, "df_energy.rds"))
df_weather_stable <- read_rds(paste0(readfile_stable, "df_weather.rds"))
df_sprt_all_stable <- read_rds(paste0(readfile_stable, "df_sprt_all.rds"))
df_seq_fs_stable <- read_rds(paste0(readfile_stable, "df_seq_fs.rds"))
df_fs_stable <- read_rds(paste0(readfile_stable, "df_fs.rds"))
df_cont_stable <- read_rds(paste0(readfile_stable, "df_cont.rds"))
df_fs_null_stable <- read_rds(paste0(readfile_stable, "df_fs_null.rds"))
df_seq_fs_null_stable <- read_rds(paste0(readfile_stable, "df_seq_fs_null.rds"))
df_model_acc_stable <- read_rds(paste0(readfile_stable, "df_model_acc.rds"))
df_fs_tmy_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy.rds"))
df_fs_tmy_null_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy_null.rds"))
df_interval_drop_stable <- read_rds(paste0(readfile_stable, "df_interval_drop.rds"))
df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))

all_sites_stable <- df_energy_stable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_stable <- df_energy_stable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_stable <- df_energy_stable %>%
  select(name) %>%
  distinct(name)

# variable dataset
readfile_variable <- str_glue("../readfiles/variable/")

df_energy_variable <- read_rds(paste0(readfile_variable, "df_energy.rds"))
df_weather_variable <- read_rds(paste0(readfile_variable, "df_weather.rds"))
df_sprt_all_variable <- read_rds(paste0(readfile_variable, "df_sprt_all.rds"))
df_seq_fs_variable <- read_rds(paste0(readfile_variable, "df_seq_fs.rds"))
df_fs_variable <- read_rds(paste0(readfile_variable, "df_fs.rds"))
df_cont_variable <- read_rds(paste0(readfile_variable, "df_cont.rds"))
df_fs_null_variable <- read_rds(paste0(readfile_variable, "df_fs_null.rds"))
df_seq_fs_null_variable <- read_rds(paste0(readfile_variable, "df_seq_fs_null.rds"))
df_model_acc_variable <- read_rds(paste0(readfile_variable, "df_model_acc.rds"))
df_fs_tmy_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy.rds"))
df_fs_tmy_null_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy_null.rds"))
df_interval_drop_variable <- read_rds(paste0(readfile_variable, "df_interval_drop.rds"))
df_interval_keep_variable <- read_rds(paste0(readfile_variable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_keep.rds"))

all_sites_variable <- df_energy_variable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_variable <- df_energy_variable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_variable <- df_energy_variable %>%
  select(name) %>%
  distinct(name)

# read functions
function_path <- "../functions/"
source(paste0(function_path, "model_fit.R"))
source(paste0(function_path, "model_pred.R"))
source(paste0(function_path, "prepost_plot.R"))

S_building <- nrow(df_fs_tmy_stable)
V_building <- nrow(df_fs_tmy_variable)
A_building <- S_building + V_building
```

# Introduction

## Conventional M&V

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. Specifically, an M&V
analyst should account for two types of baseline changes: (1) static
changes, which are known building operational changes: such as
renovation, equipment addition or removal; and (2) gradual changes,
which often occur subtly and unknown, such as the incremental adoption
of LED lighting or more efficient plug loads. Identifying and
quantifying these changes are particularly labor-intensive and
challenging. Consequently, those limitations reduce the feasibility of
M&V and complicates the quantification of estimation uncertainty, thus
impacting the accuracy and timeliness of savings assessments.

## Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we proposed a M&V method that
adopts the randomized crossover design [@raftery2024], a gold standard
from medical, agricultural and online controlled experimental (i.e. A/B
testing) studies [@blackston2019; @duan2013; @gupta2019; @mie2022;
@munro; @raseduzzaman2017]. An early application of this randomized
design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance on two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defined stopping criteria to end the M&V period early
after reaching a target level of uncertainty, among other improvements,
such as using blocked randomization by weekday. The full framework is
detailed in a previously published study [@raftery2024] along with a
case study application to a real building, and other example use cases.
In summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially shortly after the test
begins. In addition, once the desired estimate uncertainty target is
achieved, analysts can terminate the M&V study and switch to 100%
intervention. The key advantage of randomization is that if control
strategies are sampled with equal probability, the influence of other
confounding factors such as occupancy change (commonly known as one of
the non-routine events) and other more subtle long term changes in
building energy use are likely to be evenly distributed among
measurements, leading to a more accurate and unbiased assessment of the
intervention effect.

## BDG2 dataset

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly use the metered electricity and outdoor weather measurements for
running the M&V methods.

## Literature review

### Whole building approach

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline model fitting accuracy should have CV(RMSE) lower than 20%.
Thus, most research related to M&V for whole-building approach focuses
on the accuracy of baseline modeling, exploring model performance from
simple regression models to more complex machine learning techniques.
One study reviewed various models suitable for M&V applications as well
as selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.
become 'stale') over the extended period of pre- and post-analysis.

### Non-routine events impact

Non-routine events in M&V commonly refer to unexpected changes in a
building that influence a building's energy usage and are typically
unrelated to the intervention strategy. Current guidelines define
non-routine events as 'static factors' that needed to be adjusted after
projecting baseline in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies uses linear
interpolation to estimate counter-factual baseline for demand response
program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
one study points out that it is inaccurate to assume no change in the
operating conditions during the response period
[@huang_experimental_2023]. Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine events as it also
represents a change in facility operation but in a longer term.

## Objectives

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
make it even simpler by adjusting the setpoint based on outdoor weather
conditions with more description shown in Section
\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
to:

1.  Determine, using a large sample of buildings, how much more
    accurately the randomized M&V method would estimate the savings of
    the proposed intervention compared to the conventional M&V, and how
    much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset. The BDG2 dataset
    contains real-world building energy usage data collected over two
    years, inherently contains varying degrees of usage changes. By
    virtually introducing intervention effects into the existing
    dataset, we can calculate the ground-truth energy savings, which
    serve as a reference point for method comparison.

3.  Open-source implementation of the proposed randomized M&V method. We
    ensured the reproducibility of the method by making the analysis
    code open source including randomized schedule generation,
    sequential statistical analysis, energy modeling and normalized
    saving calculation. In addition, we also included code for extended
    use cases of the randomized method, such as changing sampling ratio
    and sampling intervals. Using the available open resources, building
    analysts should be able to seamlessly integrate and apply them in
    their own M&V projects.

# Method

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section.

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

knitr::include_graphics(paste0(fig_path, "flowchart.png"))
```

## Filter and clean dataset

We extracted the electricity measurements from the BDG2 dataset and
filter out all qualified buildings based on the following criteria:

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to 1.5 months of missing days.

2.  Mean electricity usage \> 0 kWh: target buildings should have active
    electricity usage.

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrenceberkeleynationallab; @mathew_big-data_2015].

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water set point reset
    control.

```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_variable$type)

stable_set <- df_energy_stable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
   left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "location")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(location, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of filtered buildings and types")
```

For all qualified buildings, we further added one more stringent
criterion to explore all buildings with very stable electricity usage
between the two years:

1.  No statistically significant difference (P-value \> 0.05) between
    the two-year electricity usage, which are sometimes assumed to be
    the case for whole-building measurement and verification.

As a result, there are only 66 buildings were labeled in this 'stable'
subset (out of total of 600 buildings) implying that the assumption that
buildings are typically stable over the timeframe involved in
conventional M&V is rarely true. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events, especially operational drifts since the
assessment last for two years. And according to the statistics shown
here, it is a typical case observed in real buildings applications.

## Apply virtue intervention

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

```

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
further consumes around 50% of the HVAC electricity. While this
assumption largely simplifies the diverse energy usage across various
building types, for the scope of this paper, we assume that 25% of the
total building electricity is used by the chilled water plant, $E_{CHW}$
[@us2012commercial]. Typically, the savings from an intervention are not
proportional to the building's hourly electricity usage, which is
generally the challenge for M&V. To address this, we mapped the
resulting electricity savings as a percentage of the plant's normal
operation, calculated as its mean electricity usage over the two-year
period. This percentage is influenced by factors such as outdoor
temperature ($OAT$), intervention supply water temperature
($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
of the day ($\delta_{occ}$, binary indicator whether during peak hours
from 9 AM to 4 PM).

$$
\text{perc_savings = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building and were applied uniformly in the filtered dataset.
Interested readers can also change the parameters in our open-source
code to simulate different scenarios.

|                        |                                                                  |       |
|------------------|------------------------------------|------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Table 1. Parameters for calculating the intervention savings.

## Run M&V methods

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post- comparison on a 12-month baseline and
12-month intervention timeline. The randomized method defines sampling
requirements as:

-   Use a daily sampling interval with the sampling time at midnight
    each day.

-   Block by day of the week with a block period of 12 weeks.

and stopping criteria as:

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but stopping
    criteria enables an early stop at the end of the satisfied blocking
    period.

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data sampled by both strategies.

-   No carryover effect using a t-test with a p-value not exceeding a
    defined significance threshold of 0.05.

-   90% confidence that energy savings exceed or do not exceed 0% using
    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
    and calculated SPRT statistics either falls below the lower
    threshold or exceeds the upper threshold.

-   Sample strategies at an equal sampling ratio (50% baseline, 50%
    intervention).

# Results

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step representing 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top for reference.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=12, fig.width=12}

#### GA ####
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff, na.rm = T)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p_accuracy <- ggarrange(p_top, p_middle, 
                        ncol = 1, nrow = 2,
                        labels = c("a)", "b)"),
                        common.legend = T, 
                        legend = "bottom")

# Timeline plot
df_time <- df_sprt_all_variable %>% 
  filter(seq != "final") %>% 
  bind_rows(df_sprt_all_stable %>% filter(seq != "final")) %>% 
  select(name, seq, n_weeks)

count <- list()
n <- 1
for (i in seq(0, 36, by = 3)){

  df_sprt <- df_time %>% 
    filter(seq == "sprt",
           n_weeks <= i)

  df_eob <- df_time %>% 
    filter(seq == "eob",
           n_weeks <= i)

  df_temp <- df_time %>% 
    filter(seq == "temp", 
           n_weeks <= i)

  count[[n]] <- tibble("n_weeks" = i, 
                       "eob" = nrow(df_eob), 
                       "temp" = nrow(df_temp), 
                       "sprt" = nrow(df_sprt))

  n <- n + 1
}

count <- bind_rows(count)

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 30, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison of satisfying all stopping criteria") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
```

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison and for clarity, we plotted conventional M&V results in grey,
randomized M&V results after satisfying all stopping criteria in green,
and the randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Acc|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using TOWT model.
As a result, the conventional M&V method exhibits an average deviation
of 6% in savings estimation, whereas the proposed randomized method
demonstrates significantly smaller deviations. If the analyst stops
immediately after satisfying all stopping criteria, the deviation is
reduced to approximately 2%. Extending the M&V period to match the
length of the conventional method (24 months) further improves accuracy
to 1%. Furthermore, by comparing subplots (a) and (b), we observe a
minor improvement in estimation accuracy when a model is fitted to the
data to account for weather differences. The model fitting is most
beneficial for the shorter dataset (randomized and stopping when all
criteria are met, typically 24-36 weeks of data), where it improves
estimation accuracy by 0.6%. This is because the weather conditions
between the first and second year for each of the locations are
relatively similar on average over this time period, and thus the effect
of adjusting for temperature as an independent variable in the model is
relatively small. Subplot (c) compares the overall timeline of the two
M&V methods. The results show that almost all buildings achieve accurate
M&V results and meet the stopping criteria within 24 weeks (for this
intervention), with the remainder meeting all stopping criteria within
36 weeks. The additional 12 weeks is usually driven by a need to span
sufficient weather conditions representative of a full year of data.
Meanwhile, the red arrow highlights the excess time required by the
conventional M&V method (24 months). Thus, the randomized M&V method
finds more accurate results in approximately a quarter of the time
required for conventional M&V. Even if it was the case that an existing
12-month baseline data was available at the start of the M&V process,
randomized M&V would still obtain a more accurate result, in less time,
than obtaining the 12-month retrofit dataset required by conventional
M&V.

We also plotted the mean error in Figure \@ref(fig:meanerr) and we
noticed that the uncertainty range associated with the conventional
method is significantly larger compared to the randomized method that
only takes 36 weeks according to subplot (c) in Figure
\@ref(fig:abstract).

```{r meanerr, fig.cap="Distribution of mean deviation from the ground-truth savings calculated by the two M&V methods", fig.height=6, fig.width=12}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Mean error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Mean error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_middle, 
          ncol = 1, nrow = 2,
          labels = c("a)", "b)")) +
  plot_annotation(title = "Mean error in savings estimation by different M&V methods")

percentiles <- df_TW_A %>%
  filter(method == "conv_diff") %>% 
  summarise(
    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
  )
```

Additionally in the literature, we found another M&V study using over
500+ commercial buildings, but their study scope was to compare the
predictive performance of a variety of M&V baseline models from simple
weekly mean to more complex machine learning methods using the
normalized mean bias error (NMBE) [@granderson_accuracy_2016]. Since we
also calculated the normalized savings estimation error, we compared our
normalized accuracy for 600+ buildings with the TOWT model and TMY
weather with their TOWT model prediction assessment in the following
table, highlighting that the model and overall building datasets are
performing similarly.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|-----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

: Table 1. Savings accuracy comparison with a similar literature

## Non-routine events impact

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=6, fig.width=12}

#### MEAN-NULL ####
rand_eob_S <- df_seq_fs_null_stable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_null_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_null_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_null_variable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_null_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_null_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                                "conv_diff" = "Conventional", 
                                "rand_eob_diff" = "Randomized", 
                                "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(mean = mean(diff)) %>% ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 6), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Mean effect in fractional savings",
       title = "Mean effect of non-routine events when using different M&V methods", 
       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
  coord_cartesian(ylim = c(-18, 18)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
the selected buildings are affected by operational drift, which is the
reason why we observe baseline electricity usage changes throughout the
two year period. To evaluate the robustness of the two M&V methods to
the effect of operational drift and non-routine events, we repeated the
M&V run on the original dataset prior to the application of the reset
control. In other words, there is no intervention applied here, so the
most reliable M&V method will be the one that detects closer to 0 kW
savings using the measurements. Therefore, the $Acc$ here are calculated
as:

$$
Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$

Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
time of satisfying all stopping criteria and at the end of a two-year
period) produced consistent results. We included the absolute deviation
($|ACC|$) figure in the supplementary material. Similarly to Figure
\@ref(fig:abstract), stopping when meeting all stopping criteria
generally yields slightly larger uncertainty than continuing for 24
months. However, it still yields a more accurate result than the
conventional method. Interestingly, similar to
[@granderson_accuracy_2016] where they obtained a median normalized
deviation around -1.25%, our analysis on conventional method shows that
the typical deviation is -1.3%, meaning that the typical building is
decreasing energy use slightly in this dataset in the absence of any
known intervention. Thus, if a building analyst were to apply an
intervention to this building and use conventional M&V to measure the
effect, they would 'detect' 1.3% energy savings even when the
intervention had no actual impact on energy performance.

# Discussion

## TOWT modeling accuracy

To evaluate modeling accuracy, we used the Coefficient of Variation of
Root-Mean Squared Error (CV(RMSE)) as the primary error metric. Since
CV(RMSE) is a normalized measure, it enables direct comparison across
different model-fitting results. According to ASHRAE Guideline 14,
whole-building baseline models should achieve a CV(RMSE) below 25% for
computing savings or 30% when less than 12 months of measurements are
available [@ashrae_ashrae_2023]. In our study, Figure \@ref(fig:towtacc)
compares the model fitting accuracy of the two M&V methods and it shows
around 87.5% of the models fitted in the randomized M&V method are code
compliant while 80% for the conventional M&V method. As the quantity of
the training set is the same, this highlights the impact of data
sampling on model fitting accuracy, and for TOWT, the distribution of
outdoor weather conditions sampled in the training dataset is
influential. The conventional method trains the model using the entire
first-year dataset, if the location is heating or cooling dominant, the
data-driven model is trained using colder or warmer outdoor conditions
leading to biased model coefficients. Although the random sampling in
our study does not specifically block the outdoor temperature factor,
the CV(RMSE) distribution shown in the figure indicates a significant
improvement in model fitting accuracy.

```{r towtacc, fig.cap="TOWT model fitting accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}

cvrmse <- df_model_acc_stable %>% 
  bind_rows(df_model_acc_variable) %>% 
  pivot_longer(c(conv, rand), 
               names_to = "method", 
               values_to = "cvrmse") %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                              "rand" = "Randomized\n(24 months)", 
                              "conv" = "Conventional"))

result <- t.test(filter(cvrmse, method == "Randomized\n(24 months)")$cvrmse, 
                 filter(cvrmse, method == "Conventional")$cvrmse, 
                 exact = TRUE,
                 alternative = "less",
                 mu = 0,
                 conf.int = TRUE, 
                 conf.level = 0.90,
                 var.equal = TRUE)$p.value

cvrmse %>% 
  ggplot(aes(x = method, y = cvrmse, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  annotate(geom = "text", 
           x = 1.5, 
           y = 40, 
           size = 4, 
           label = paste0("P value = ", round(result, digits = 3))) +
  geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
  geom_text(data = . %>% 
              group_by(method) %>% 
              summarise(mean = mean(cvrmse, na.rm = T)) %>% 
              ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
  geom_hline(yintercept = 25, color = "red", lty = "dashed") +
  annotate(geom = "text", 
           x = 0.5, 
           y = 27, 
           size = 4, 
           color = "red", 
           label = paste0(25, " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %")) +
  coord_cartesian(ylim = c(0, 50)) +
  labs(x = NULL, 
       y = NULL, 
       color = NULL, 
       title = "TOWT model fitting accuracy by different M&V methods", 
       subtitle = "CV(RMSE)") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Sampling interval impact

In some cases, particularly those involving large thermal mass and
active thermal energy storage systems or operating continuously, there
can be a substantial time-lagged effect associated with interventions.
This means that the effect of one control strategy may persist after
switching to another. To mitigate such carryover effects, building
analysts may choose to drop non-consecutive days in the switchback
experiment. For example, consider a heat pump hot water heater where the
tank is charged at very different times of the day by the intervention
strategy compared to the baseline strategy. When switching strategy, the
tank may be much warmer or cooler than typical for that strategy, thus
influencing subsequent measurements, particularly the performance
measured on the first day (i.e., a non-consecutive day). In such cases,
increasing the sampling interval (e.g., sample strategies every 2 or 3
days instead of daily) and discarding measurements from those
non-consecutive days can eliminate carryover. However, this approach
reduces the total data collected: for instance, sampling every 3 days
would drop 1/6 of measurements and sampling every 2 day means would
remove 1/4 of measurements. Therefore, by increasing the sampling
interval, there is an accuracy penalty due to a reduction in how often
samples occur, and when decreasing the sampling interval, there is an
accuracy penalty from dropping more non-consecutive days. To quantify
such trade-off, \@ref(fig:absinterval) shows the accuracy of different
sampling intervals and the consequence of dropping the non-consecutive
days in the dataset at the time when randomized M&V produces a result
(after satisfying all stopping criteria).

```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}

#### ABS-INT ####
df_seq_interval_fs_keep_stable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_seq_fs_stable %>% 
              filter(seq == "eob") %>% 
              select(-seq) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "fs"))

df_seq_interval_fs_keep_variable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_seq_fs_variable %>% 
              filter(seq == "eob") %>% 
              select(-seq) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "fs"))

df_drop <- df_seq_interval_fs_drop_stable %>% 
  bind_rows(df_seq_interval_fs_drop_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "drop")

df_keep <- df_seq_interval_fs_keep_stable %>% 
  bind_rows(df_seq_interval_fs_keep_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "keep")

df_MW_A <- df_drop %>% 
  bind_rows(df_keep) %>% 
  left_join(rbind(df_fs_stable %>% 
                    filter(scenario == "ref" & method == "true"), 
                  df_fs_variable %>% 
                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(diff = abs(savings - eob)) %>% 
  filter(is.finite(diff))

p_top <- df_MW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_jitter(alpha = 0.8, 
              size = 0.5, 
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(group, interval) %>% 
              summarise(mean = mean(diff, inf.rm = T)) %>% 
              ungroup(), 
            aes(x = interval, 
                y = mean, group = group,
                label = paste0(round(mean, digits = 1), " %")), 
            color = "white",
            position = position_dodge(width = 0.75)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with no model normalization on weather")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# with tmy version
df_seq_interval_nm_keep_stable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_sprt_all_stable %>% 
              filter(seq == "eob") %>% 
              select(-c(seq, n_weeks)) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "annual"))

df_seq_interval_nm_keep_variable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_sprt_all_variable %>% 
              filter(seq == "eob") %>% 
              select(-c(seq, n_weeks)) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "annual"))

df_drop <- df_seq_interval_nm_drop_stable %>% 
  bind_rows(df_seq_interval_nm_drop_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "drop")

df_keep <- df_seq_interval_nm_keep_stable %>% 
  bind_rows(df_seq_interval_nm_keep_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "keep")

df_TW_A <- df_drop %>% 
  bind_rows(df_keep) %>% 
  left_join(rbind(df_fs_stable %>% 
                    filter(scenario == "ref" & method == "true"), 
                  df_fs_variable %>% 
                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(diff = abs(savings - eob)) %>% 
  filter(is.finite(diff))

p_bottom <- df_TW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_jitter(alpha = 0.8, 
              size = 0.5, 
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(group, interval) %>% 
              summarise(mean = mean(diff, na.rm = T)) %>% 
              ungroup(), 
            aes(x = interval, 
                y = mean, 
                group = group,
                label = paste0(round(mean, digits = 1), " %")),
            color = "white",
            position = position_dodge(width = 0.75)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model normalized on TMY weather")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_bottom, 
          ncol = 1, nrow = 2, 
          align = "hv", 
          labels = c("a)", "b)"),
          common.legend = T, 
          legend = "bottom") +
  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")

```

To quantify the impact of reduced randomness, all measurements are kept
when calculating the distribution shown in the darker-colored set. The
results suggest a gradual increase in error from 2.4% to 3.7% when the
sampling interval increases from daily to weekly intervals. This is due
to the fact that increasing the sample interval means reducing
randomization. To quantify the combined impact of data loss due to
preventing the carryover effect, we generated the lighter-colored set,
which drops all non-consecutive days. As a result, we noticed a
significant increase in estimation error when sampling daily, which led
to a similar accuracy compared to weekly intervals. Therefore,
considering this trade-off, we recommend using a two-day sampling
interval if the carryover effect is likely but expected to last less
than one day, and then normalizing the estimation via energy modeling.
We also included the mean deviation distribution and the accuracy
distribution calculated after all stopping criteria were satisfied in
the supplementary material and the results show the same indication.

```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=6, fig.width=12}
# timeline to finish sequential test
df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
  mutate(group = "drop", 
         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
  select(name, seq, n_weeks, interval, group)

interval_1_keep <- bind_rows(df_sprt_all_stable, df_sprt_all_variable) %>%
  select(c(name, seq, n_weeks)) %>%
  filter(seq != "final") %>%
  mutate(seq = as.factor(seq)) %>%
  mutate(interval = 1, 
         group = "keep")

df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_variable) %>%
  mutate(group = "keep", 
         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
  select(name, seq, n_weeks, interval, group) %>% 
  bind_rows(interval_1_keep)

df_MW_A <- bind_rows(df_keep, df_drop) %>% 
  filter(seq != "final") %>% 
  filter(seq == "eob") %>% 
  mutate(seq = as.factor(seq), 
         seq = recode_factor(seq, 
                             "eob" = "Buildings finishing all criteria"), 
         interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept"), 
         n_weeks = as.factor(n_weeks), 
         n_weeks = recode_factor(n_weeks, 
                                 "12" = "12\nweeks", 
                                 "24" = "24\nweeks", 
                                 "36" = "36\nweeks", 
                                 "48" = "48\nweeks")) %>% 
  group_by(n_weeks, group, interval) %>% 
  summarise(n = n()) %>%
  ungroup()

levA <- levels(df_MW_A$n_weeks)
levB <- levels(df_MW_A$group)
levC <- levels(df_MW_A$interval)

# 2. Expand to all combinations of factorA and factorB
all_combinations <- expand_grid(
  n_weeks = levA,
  group = levB, 
  interval = levC, 
)

df_MW_A <- all_combinations %>%
  left_join(df_MW_A, by = c("n_weeks", "group", "interval")) %>% 
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  group_by(group, interval) %>% 
  mutate(n = ifelse(n > sum(n) * 0.01, n, 0), 
         perc = n / sum(n) * 100) %>% 
  ungroup()

df_MW_A %>% 
  ggplot(aes(x = n_weeks, y = perc, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~interval, nrow = 1) +
  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), position = position_dodge(1)) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Percentage of buildings using randomized M&V when satisfying all stopping critiera", 
       subtitle = "Counted at the end of each 12-week blocking period") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

Sampling at different intervals and excluding non-consecutive days also
impact other stopping criteria, such as the time needed to satisfy
statistical uncertainty and weather conditions thresholds. In Figure
\@ref(fig:intervaltime), we compared the percentage of buildings
satisfying all stopping criteria under different sampling intervals,
with or without dropping non-consecutive days. The results indicate that
randomized M&V concludes more quickly when all data are retained,
particularly at shorter intervals. For instance, when sampling at a
one-day interval, 90% of buildings meet all criteria within 24 weeks if
no data are dropped, but this number decreases significantly to 52% when
non-consecutive days (account for 1/3 of the data) are excluded. As the
sampling interval increases, this difference gradually diminishes: when
sampling at a weekly interval, removing non-consecutive days has no
impact on the timeline of meeting stopping criteria.

## Sampling ratio impact

Another advantage of using randomized M&V is the flexibility of changing
the sampling ratio after the target savings are detected. For instance,
the building owner can continue sampling at a 50%/50% ratio between the
baseline and the intervention to further reduce the uncertainty
associated with the savings. Alternatively, they could switch to 100%
intervention to maximize cost savings, though this approach risks the
baseline becoming outdated over time, which is another trade-off to
consider. A middle-ground approach may be to continue sampling at a
20%/80% ratio or 10%/90% between the baseline and the intervention. To
compare those possible choices to a building owner, figure
\@ref(fig:abscont) shows the results of changing the original 50%/50%
sampling ratio after satisfying all stopping criteria and continuing a
new ratio till the end of the two-year period. Those results were
calculated without dropping non-consecutive days and include the initial
50%/50% sampled data. On the far left, we also provided the case when
the building owner continues with no change in sampling ratio. The
results show that as the sampled baseline days and intervention days
become unbalanced, the savings estimation error grows from 1% to 5%.
This is reasonable since sampling at 10%/90% is similar to the
conventional M&V method. And on the far right, we show again the
accuracy of the conventional M&V method, which is still worse then
sampling at 10%/90%. Also, We include the mean deviation version in the
supplementary material.

```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months", fig.height=6, fig.width=12}

#### ABS-CONT ####
conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         ratio = "conv") %>% 
  select(name, ratio, diff)

rand_cont_S <- df_cont_stable %>% 
  select(-cont_tmy) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs)) %>% 
  select(name, diff, ratio)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         ratio = 0) %>% 
  select(name, diff, ratio)

df_MW_S <- rbind(rand_cont_S, rand_final_S, conv_S)

# plot for the variable subset
conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         ratio = "conv") %>% 
  select(name, ratio, diff)

rand_cont_V <- df_cont_variable %>% 
  select(-cont_tmy) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs)) %>% 
  select(name, diff, ratio)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         ratio = 0) %>% 
  select(name, diff, ratio)

df_MW_V <- rbind(rand_cont_V, rand_final_V, conv_V)

df_MW_A <- rbind(df_MW_S, df_MW_V)

df_MW_A %>% 
  mutate(ratio = as.factor(ratio), 
         ratio = recode_factor(ratio, "0" = "Randomized\n(50/50)", "1" = "Randomized\n(20/80)", "2" = "Randomized\n(10/90)", "conv" = "Conventional")) %>% 
  ggplot(aes(x = ratio, y = diff, fill = ratio)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(ratio) %>% 
              summarise(mean = mean(diff)) %>% 
              ungroup(), 
            aes(x = ratio, y = mean, label = paste0(round(mean, digits = 1), " %"))) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       title = "Absolute error in savings estimation fo randomized M&V using different sampling ratio",
       subtitle = str_glue("All {A_building} buildings with with no model normalization on weather")) +
  coord_cartesian(ylim = c(0, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Limitations

Aside from the general limitation that this method only applies to the
subset of interventions in buildings that are easily switchable, we
identify two key limitations in this study:

1.  The simulated intervention remains somewhat generic and simplistic
    considering the diverse building types and climate zones in the BDG2
    dataset. As the primary focus is to accurately detect the
    intervention (or any intervention) effect so for simplicity, we
    applied the parameters listed in Table 1 uniformly across all
    buildings. Yet, in some cases, raising the water temperature by 1°C
    might yield more or less than the specified 8% electricity savings
    due to target buildings' demand flexibility. For example, one field
    validation of HVAC optimization shows additional savings might be
    limited if other best practice controls are already implemented
    [@granderson2018].
2.  For the main analysis, we assumed no carryover effect and a daily
    sampling interval would suffice for most commercial buildings, but
    certainly exceptions exist. However, in the discussion, we addressed
    such concern by comparing different sampling strategies and the
    impact of dropping measurements from non-consecutive days.

# Conclusion

This research demonstrated the application of a randomized
whole-building measurement and verification (M&V) method, comparing its
performance to the conventional approach described in the IPMVP and in
ASHRAE Guideline 14, using a large, open-source commercial building
dataset. The porposed M&V method leverages the randomized experimental
design concept from other scientific fields, along with statistical
sequential inference techniques, to determine when target savings are
detected. We used a virtual control retrofit case---resetting the
chilled water setpoint based on outdoor weather conditions---and applied
it to over 600 filtered commercial buildings. By comparing the savings
estimations of the conventional method with the novel randomized method,
we found that the randomized approach provides faster and more robust
savings estimations.

Specifically, we showed that throughout 11 different locations assessed
in this study, the randomized M&V can provide a saving estimation by 36
weeks (with the majority finished by 24 weeks) once all stopping
criteria are satisfied. In contrast, the conventional method requires a
full range of baseline and intervention measurements under normal
operating conditions, typically taking 6-9 months for each phase. Most
importantly, we verified that with a reduced M&V timeline, the
randomized method can estimate savings more accurately by showing that
the absolute error is only 1 - 2 % for a typical building. Whereas for
the much longer two-year conventional method, the estimation error was
approximately 6% in a typical building. We also evaluated the impact of
non-routine events especially more gradual and subtle operational drift
on the proposed M&V method and the results show that baseline changes in
the post-retrofit period can deviate savings estimated using the
conventional method. As a comparison, we found those events have a very
negligible impact on the savings estimated using the randomized method,
demonstrating robustness to this issue. Lastly, we consider scenarios
where there is a known carryover effect from switching between
strategies in the building and assess the impact of dropping samples
where the strategy is non-consecutive and varying the sampling interval.
Through that process, we showed that for the typical building, if a
carryover is likely to be present, then using a 3-day sampling interval
and dropping non-consecutive samples yields the optimal design.

We also discussed the impact of model fitting accuracy when using the
two methods and we found a significant improvement in our method due to
random sampling. We also provide a detailed assessment of using
different sampling ratios, those results provide useful considerations
to building analysts when generating randomized schedules for a variety
of use cases.

# CRediT authorship contribution statement

**Aoyu Zou**: Conceptualization, Data curation, Formal analysis,
Methodology, Investigation, Software, Writing - original draft,
Writing - review & editing. **Paul Raftery**: Conceptualization, Funding
acquisition, Formal analysis, Methodology, Investigation, Project
administration, Supervision, Writing - review & editing. **Stefano
Schiavon**: Conceptualization, Methodology, Investigation, Supervision,
Writing - review & editing. **Carlos Duarte**: Methodology,
Investigation, Supervision, Writing - review & editing. **Gail Brager**:
Supervision, Writing - review & editing.

# Reproducibility

A reproducible example with analysis code is available (MIT license) at
<https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.

# Declaration of competing interest

The authors declare that they have no known competing financial
interests or personal relationships that could have appeared to
influence the work reported in this paper.

# Acknowledgements

# References