### 2025-06-27 code/manuscript.Rmd
updated supplementary material and manuscript

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 2 years). Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../paper"
    )
  })
---

```{r setup, include = FALSE, cache = FALSE}

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      # cache = T,
                      dpi = 300,
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/manuscript/figs/")

# str(knitr::opts_chunk$get()) # see for all options

```

```{r prep, include = FALSE, cache = FALSE}
#### LIBRARIES ####
require(pacman)

# load packages using pacman
pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, gridExtra, patchwork, RColorBrewer)

# turn off scientific notation
options(scipen = 999, digits = 15)

# set directory
here::i_am("manuscript.rmd")

# set default theme for ggplot
theme_set(theme_minimal())

# define base ggplot theme
theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
             plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
             plot.background = element_rect(fill = "white", colour = NA),
             panel.grid.minor = element_blank(),
             panel.grid.major = element_blank(),
             axis.text = element_text(size = 10),
             strip.text = element_text(size = 10, color = "grey20", face = "bold"),
             strip.background = element_blank())

# colors
ls_colors <- c("Baseline" = "grey10",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#6baed6",
               "Randomized\n(50/50)" = "#6baed6",
               "Randomized\n(20/80)" = "#6baed6", 
               "Randomized\n(10/90)" = "#6baed6", 
               "Continued\n(20/80)" = "#c6dbef", 
               "Continued\n(10/90)" = "#c6dbef",
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Dropped (24 months)" = "#4292c6",
               "Kept (24 months)" = "#08519c",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

# parameters
ctr_params <- list(peak_hours = 10:16,
                   chwl_perc = 0.25,
                   step_perc = 0.08,
                   conv_swt = 6,
                   weather_knots = c(15, 25),
                   swt_knots = c(12, 7),
                   coe_peak = 0.8,
                   coe_off = 1.2,
                   enable_temp = 8)

occ_params <- list(change_start = 5,
                   change_end = 8,
                   change = 20)

#### FUNCTIONS ####
# Function defined to read downloaded tmy files
get_tmy <- function(all_sites, readfile_path){

  df_tmy <- data.frame()

  for (site in all_sites){
    df <- read_csv(paste0(readfile_path, "tmy/", str_glue("{site}.epw")),
                   skip = 8, col_types = "ddddd-d---------------------------------",
                   col_names = c("year", "month", "day", "hour", "min", "tmy")) %>%
      mutate(year = 2017,
             time = ymd_h(paste(paste(year, month, day, sep = "-"), hour, sep = " ")),
             temp = tmy) %>%
      dplyr::select(time, temp) %>% 
      mutate(site = site)

    df_tmy <- rbind(df_tmy, df)
  }

  return(df_tmy)
}

# Function defined to add chwst reset intervention
run_reset <- function(df_baseline){

  mean <- mean(df_baseline$base_eload, na.rm = T) * ctr_params$chwl_perc

  grad <- (ctr_params$swt_knots[2] - ctr_params$swt_knots[1]) / 
    (ctr_params$weather_knots[2] - ctr_params$weather_knots[1])

  interc <- ctr_params$swt_knots[2] - (ctr_params$weather_knots[2] * grad)

  df_interv <- df_baseline %>% 
    mutate(swt = t_out * grad + interc, 
           chwl = mean,
           hour = hour(datetime)) %>% 
    mutate(swt = ifelse(swt > ctr_params$swt_knots[1], ctr_params$swt_knots[1], ifelse(swt < ctr_params$swt_knots[2], ctr_params$swt_knots[2], swt)), 
           temp_savings = ifelse(t_out >= ctr_params$enable_temp, (swt - ctr_params$conv_swt) * ctr_params$step_perc, 0), 
           time_adj = ifelse(hour %in% ctr_params$peak_hours, ctr_params$coe_peak, ctr_params$coe_off), 
           perc_savings = temp_savings * time_adj, 
           savings = chwl * perc_savings, 
           interv_eload = base_eload - savings) %>% 
    select(datetime, base_eload, interv_eload, t_out)

  return(df_interv)
}

# Function defined to interpolate NAs
run_interpo <- function(df_all){

  na_counts <- df_all %>%
    mutate(date = date(timestamp)) %>%
    group_by(date) %>%
    summarize(na_hours = sum(is.na(eload)))

  # Filter out days with more than half of the hours having NAs
  valid_days <- na_counts %>%
    filter(na_hours <= 12) %>%
    pull(date)

  df_filtered <- df_all %>%
    filter(date(timestamp) %in% valid_days)

  df_filtered <- df_filtered %>%
    mutate(across(c(eload, t_out), ~ zoo::na.approx(., na.rm = FALSE))) %>% 
    rename(datetime = timestamp, 
           base_eload = eload)

  return(df_filtered)
}

# Function defined to adjust the plot scale
get_scale <- function(eload, range = 2){

  min_y <- mean(eload, na.rm = T) - range * sd(eload, na.rm = T)
  max_y <- mean(eload, na.rm = T) + range * sd(eload, na.rm = T)

  return(c(min_y, max_y))
}
```

```{r readdata, include=FALSE}

#### READ DATA ####
df_loc <- read.csv("../readfiles/loc_map.csv")

# stable dataset
readfile_stable <- str_glue("../readfiles/stable/")
fig_path = "../figs/manuscript/"

df_energy_stable <- read_rds(paste0(readfile_stable, "df_energy.rds"))
df_weather_stable <- read_rds(paste0(readfile_stable, "df_weather.rds"))
df_sprt_all_stable <- read_rds(paste0(readfile_stable, "df_sprt_all.rds"))
df_seq_fs_stable <- read_rds(paste0(readfile_stable, "df_seq_fs.rds"))
df_fs_stable <- read_rds(paste0(readfile_stable, "df_fs.rds"))
df_cont_stable <- read_rds(paste0(readfile_stable, "df_cont.rds"))
df_fs_null_stable <- read_rds(paste0(readfile_stable, "df_fs_null.rds"))
df_seq_fs_null_stable <- read_rds(paste0(readfile_stable, "df_seq_fs_null.rds"))
df_model_acc_stable <- read_rds(paste0(readfile_stable, "df_model_acc.rds"))
df_fs_tmy_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy.rds"))
df_fs_tmy_null_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy_null.rds"))
df_interval_drop_stable <- read_rds(paste0(readfile_stable, "df_interval_drop.rds"))
df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))

all_sites_stable <- df_energy_stable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_stable <- df_energy_stable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_stable <- df_energy_stable %>%
  select(name) %>%
  distinct(name)

# variable dataset
readfile_variable <- str_glue("../readfiles/variable/")

df_energy_variable <- read_rds(paste0(readfile_variable, "df_energy.rds"))
df_weather_variable <- read_rds(paste0(readfile_variable, "df_weather.rds"))
df_sprt_all_variable <- read_rds(paste0(readfile_variable, "df_sprt_all.rds"))
df_seq_fs_variable <- read_rds(paste0(readfile_variable, "df_seq_fs.rds"))
df_fs_variable <- read_rds(paste0(readfile_variable, "df_fs.rds"))
df_cont_variable <- read_rds(paste0(readfile_variable, "df_cont.rds"))
df_fs_null_variable <- read_rds(paste0(readfile_variable, "df_fs_null.rds"))
df_seq_fs_null_variable <- read_rds(paste0(readfile_variable, "df_seq_fs_null.rds"))
df_model_acc_variable <- read_rds(paste0(readfile_variable, "df_model_acc.rds"))
df_fs_tmy_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy.rds"))
df_fs_tmy_null_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy_null.rds"))
df_interval_drop_variable <- read_rds(paste0(readfile_variable, "df_interval_drop.rds"))
df_interval_keep_variable <- read_rds(paste0(readfile_variable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_keep.rds"))

all_sites_variable <- df_energy_variable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_variable <- df_energy_variable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_variable <- df_energy_variable %>%
  select(name) %>%
  distinct(name)

# read functions
function_path <- "../functions/"
source(paste0(function_path, "model_fit.R"))
source(paste0(function_path, "model_pred.R"))
source(paste0(function_path, "prepost_plot.R"))

S_building <- nrow(df_fs_tmy_stable)
V_building <- nrow(df_fs_tmy_variable)
A_building <- S_building + V_building
```

### Graphical abstract {.unnumbered}

```{r graphicalabstract, out.width = "80%"}

knitr::include_graphics("../figs/manuscript/graphical_abs.jpg", dpi = 900)
```

### Keywords {.unnumbered}

Measurement and verification (M&V); Controls; Randomized switchback
design; Uncertainty quantification; Building energy analysis; Heating
ventilation and air conditioning

### Highlights {.unnumbered}

-   Verified the improvements of the randomized switchback method using
    a large open-source dataset
-   Addressed the influence of non-routine events and also more subtle
    long-term 'operational drift' in buildings
-   Investigated the trade-off of selecting longer sampling intervals to
    account for carryover effect and weather dependency of the method
-   Discussed the impact of selecting different sampling ratio on the
    M&V accuracy

# Introduction

## Conventional M&V

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level, where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. In conventional practice, the process
begins with baseline measurements taken during the year before
implementing the energy-efficiency retrofit, followed by the same
measurement procedure during the year after the intervention. After
collecting two years of data, an M&V analyst fits an energy prediction
model, using variables such as outdoor temperature and time
[@mathieu_quantifying_2011] to project baseline energy consumption in
the post-retrofit period. The difference between the counter-factual
baseline and the measured intervention period energy consumption
represents the energy savings. One of the key drawbacks of this method
is its reliance on a two-year timeline to quantify savings, during which
baseline measurements can become outdated due to changes in building
performance caused by non-routine events unrelated to the intervention.
Specifically, an M&V analyst should account for two types of baseline
changes: (1) static changes, which are known building operational
changes: such as renovation, equipment addition or removal, and changes
related to tenant turnover; and (2) gradual changes, which often occur
subtly and unknown, such as the incremental adoption of LED lighting or
more efficient plug loads. Identifying and quantifying these changes are
particularly labor-intensive and challenging. Consequently, those
limitations reduce the feasibility of M&V and complicate the
quantification of estimation uncertainty, thus impacting the accuracy
and timeliness of savings assessments.

## Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. An early application of this
randomized design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance of two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defining stopping criteria to end the M&V period early
after reaching a target level of uncertainty, using blocked
randomization by weekday. The full framework is detailed in a previously
published study [@raftery2024] along with a case study application to a
real building, and other example use cases. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week and seasons. For example, given a
10-week M&V period for one intervention, the balanced randomized
schedule would equally sample five Mondays with the baseline strategy in
operation and five Mondays with the intervention strategy in operation.
The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off, which is a small subset of all the interventions performed
in buildings. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially shortly after the test begins. In
addition, once the desired estimate uncertainty target is achieved,
analysts can terminate the M&V study and switch to 100% intervention.
The key advantage of randomization is that if control strategies are
sampled with equal probability, the influence of other confounding
factors such as occupancy change (a common non-routine event) and other
more subtle long term changes in building energy use are likely to be
evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention effect.

## BDG2 dataset

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies, acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly used the metered electricity and outdoor weather measurements for
evaluating the M&V methods.

## Literature review

### Whole building approach

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have CV(RMSE) lower than
25% for 12-month data collection. Thus, most research related to M&V for
the whole-building approach focuses on the accuracy of baseline
modeling, exploring model performance from simple regression models to
more complex machine learning techniques. One study reviewed various
models suitable for M&V applications as well as selected input features
[@alrobaie_review_2022] and another study provided a definitive
methodology to apply machine learning models for M&V use cases
[@gallagher_development_2018]. In addition, a few studies investigated
the critical performance metrics to evaluate the developed baseline
models [@granderson_automated_2015; @granderson_development_2014] and
compared a variety of models using those metrics
[@granderson_accuracy_2016]. These studies made significant
contributions by emphasizing the uncertainty associated with the
model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.,
become 'stale') over the extended period of pre- and post-analysis.

### Changes in building baseline measurements

Changes in a building that influence its energy usage and are unrelated
to the intervention being studied by the M&V method are commonly
referred to as non-routine events. Current guidelines define non-routine
events as 'static factors' that need to be adjusted after projecting the
baseline in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE Guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments, such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies use linear
interpolation to estimate counter-factual baseline for the demand
response program [@beil_round-trip_2015; @keskar_commercial_2020].
Consequently, one study points out that it is inaccurate to assume no
change in the operating conditions during the response period (Huang,
Katipamula, and Lutes 2023). Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. We found a similar concept briefly defined in
Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during multi-year
M&V project. However, currently there is no detailed instruction
provided. If M&V analysts are unaware of such changes and lack an
appropriate adjustment method (e.g., replacing filters before the
intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine event, as it also
represents a change in facility operation that occurs gradually over a
longer period of time.

## Objectives

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
made it even simpler by adjusting the setpoint based on outdoor weather
conditions (see description in Section
\@ref(apply-virtue-intervention)). By conducting such an analysis, we
aim to:

1.  Determine, using a large sample of buildings, how much more
    accurately the randomized M&V method would estimate the savings of
    the proposed intervention compared to the conventional M&V, and how
    much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset. The BDG2 dataset
    contains real-world building energy usage data collected over two
    years, which inherently contains varying degrees of usage changes.
    By virtually introducing intervention effects into the existing
    dataset, we can calculate the ground-truth energy savings, which
    serve as a reference point for method comparison.

3.  Open-source implementation of the proposed randomized M&V method. We
    ensured the reproducibility of the method by making the analysis
    code open source, including randomized schedule generation,
    sequential statistical analysis, energy modeling and normalized
    saving calculation. In addition, we also included code for extended
    use cases of the randomized method, such as changing sampling ratio
    and sampling intervals. Using the available open resources, building
    analysts should be able to seamlessly integrate and apply them in
    their own M&V projects.

# Method

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section.

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="70%"}

knitr::include_graphics(paste0(fig_path, "flowchart.png"))
```

## Filter and clean dataset

We extracted the electricity measurements from the BDG2 dataset and
filter out all qualified buildings based on the following criteria:

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to 1.5 months of missing days.

2.  Mean electricity usage \> 0 kWh: target buildings should have active
    electricity usage.

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrence_berkeley_national_laboratory_buildings_2015;
    @mathew_big-data_2015].

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water set point reset
    control.

```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_variable$type)

stable_set <- df_energy_stable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
   left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "location")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(location, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of filtered buildings and types")
```

For all qualified buildings, we also create a feature to identify
buildings with very stable electricity usage overall by identifying
those with no statistically significant difference (P-value \> 0.05)
between their two-year electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings. While we did not adjust for weather
differences between the two years when generating this high-level
feature, this result highlights that the assumption that buildings are
typically stable over the timeframe involved in conventional M&V methods
is rarely true. As mentioned in \@ref(non-routine-events-impact), such
variability is largely associated with non-routine events, especially
operational drifts since the assessment, even under ideal situations,
lasts for two years. And according to the statistics shown here, it is a
typical case observed in real buildings applications.

## Apply virtue intervention

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that resets the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

```

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
further consumes around 50% of the HVAC electricity. While this
assumption largely simplifies the diverse energy usage across various
building types, for the scope of this paper, we assume that 25% of the
total building electricity is used by the chilled water plant, $E_{CHW}$
[@us2012commercial]. Typically, the savings from an intervention are not
proportional to the building's hourly electricity usage, which is
generally the challenge for M&V. To address this, we mapped the
resulting electricity savings as a percentage of the plant's normal
operation, calculated as its mean electricity usage over the two-year
period. This percentage is influenced by factors such as outdoor
temperature ($OAT$), intervention supply water temperature
($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
of the day ($\delta_{occ}$, binary indicator whether during peak hours
from 9 AM to 4 PM).

$$
\text{perc_savings = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building and were applied uniformly in the filtered dataset.
Interested readers can also change the parameters in our open-source
code to simulate different scenarios.

| Parameter              | Description                                                      | Value |
|-------------------|-----------------------------------|-------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Table 1. Parameters for calculating the intervention savings.

## Run M&V methods

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post-comparison on a 12-month baseline and 12-month
intervention timeline. The randomized method defines sampling
requirements as:

-   Use a daily sampling interval with the sampling time at midnight
    each day.

-   Block by day of the week with a block period of 12 weeks.

and stopping criteria as:

-   A minimum and maximum of 12 and 108 weeks, respectively. The
    randomized schedule covers the entire two-year period but the
    stopping criteria enable an early stop at the end of the satisfied
    blocking period.

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data is sampled by both strategies.

-   90% confidence that energy savings exceed or do not exceed 0% using
    the sequential probability ratio test (SPRT). Medium effect size (d
    = 0.5) quantified by Cohen's d and calculated SPRT statistics either
    fall below the lower threshold or exceed the upper threshold.

-   Sample strategies at an equal sampling ratio (50% baseline, 50%
    intervention).

# Results

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage ($E_{intervention}$ is the
'true' electricity usage calculated from Figure \@ref(fig:chwst) and
above equations) and $E'$ indicates the estimated electricity usage
either through the conventional method or the new randomized method. We
present the distribution of savings estimation accuracy using boxen
plots, also known as letter plots, which is an advanced variation of the
box plot designed to extend beyond the interquartile range (IQR) by
progressively dividing the data into smaller percentiles, revealing more
detail in the tails of the distribution. We set the division parameter
to k = 4, meaning the entire box area represents the data distribution
from 6.25% to 93.75%. As the steps move closer to the center line (50%
median), the distribution range progressively narrows with the next step
representing 12.5% to 87.5%, followed by 25% to 75%. In addition, for
each boxen plot, we also showed the mean value of the distribution on
top for reference.

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison and for clarity, we plotted conventional M&V results in grey,
randomized M&V results after satisfying all stopping criteria in green,
and the randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Acc|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using a
time-of-week-temperature (TOWT) model [@mathieu_quantifying_2011]. As a
result, the conventional M&V method exhibits a median deviation of 5% in
savings estimation, whereas the randomized method demonstrates
significantly smaller deviations. If the analyst stops immediately after
satisfying all stopping criteria, the deviation is reduced to
approximately 1.5%. Extending the M&V period to match the same time
period of the conventional method (24 months) further improves accuracy
to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
observed a minor improvement in estimation accuracy when a model is
fitted to the data to account for weather differences. The model fitting
is most beneficial for the shorter dataset (randomized and stopping when
all criteria are met, typically 24-36 weeks of data), where it reduces
deviation from 1.5% to 1.2% (around 20% improvement). This is because
the weather conditions between the first and second year for each of the
locations are relatively similar on average over this time period, and
thus the effect of adjusting for temperature as an independent variable
in the model is relatively small. We found 83% of all filtered buildings
show less error when using the randomized M&V, and by comparing the
error distribution in subplot (b), we noticed the medium deviation drops
from 4.6% to 1.2% (around 74% improvement) at the time of satisfying all
stopping criteria. Subplot (c) compares the overall timeline of the two
M&V methods. The results show that almost all buildings achieve accurate
M&V results and meet the stopping criteria within 24 weeks (for this
intervention), with the remainder meeting all stopping criteria within
36 weeks. The additional 12 weeks are usually driven by a need to span
sufficient weather conditions representative of a full year of data.
Meanwhile, the red arrow highlights the excess time required by the
conventional M&V method (24 months). Thus, the randomized M&V method
finds more accurate results in approximately a quarter of the time
required for conventional M&V. Even if it was the case that an existing
12-month baseline data was available at the start of the M&V process,
randomized M&V would still obtain a more accurate result, in less time,
than obtaining the 12-month retrofit dataset required by conventional
M&V.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}

#### GA ####
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p_accuracy <- ggarrange(p_top, p_middle, 
                        ncol = 1, nrow = 2,
                        labels = c("a)", "b)"),
                        common.legend = T, 
                        legend = "bottom")

# Timeline plot
df_time <- df_sprt_all_variable %>% 
  filter(seq != "final") %>% 
  bind_rows(df_sprt_all_stable %>% filter(seq != "final")) %>% 
  select(name, seq, n_weeks)

count <- list()
n <- 1
for (i in seq(0, 36, by = 3)){

  df_sprt <- df_time %>% 
    filter(seq == "sprt",
           n_weeks <= i)

  df_eob <- df_time %>% 
    filter(seq == "eob",
           n_weeks <= i)

  df_temp <- df_time %>% 
    filter(seq == "temp", 
           n_weeks <= i)

  count[[n]] <- tibble("n_weeks" = i, 
                       "eob" = nrow(df_eob), 
                       "temp" = nrow(df_temp), 
                       "sprt" = nrow(df_sprt))

  n <- n + 1
}

count <- bind_rows(count)

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 30, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")

MW_summary <- df_MW_A %>% 
  pivot_wider(names_from = "method", values_from = "diff") %>% 
  group_by(name) %>% 
  summarise(improv_eob = (conv_diff - rand_eob_diff), 
            improv_final = (conv_diff - rand_final_diff)) %>% 
  ungroup() %>% 
  summarise(across(c(improv_eob, improv_final), median))

TW_summary <- df_TW_A %>% 
  pivot_wider(names_from = "method", values_from = "diff") %>% 
  group_by(name) %>% 
  summarise(improv_eob = (conv_diff - rand_eob_diff), 
            improv_final = (conv_diff - rand_final_diff), 
            improv = ifelse(improv_eob >= 0, 1, 0)) %>% 
  ungroup() %>% 
  mutate(perc = sum(improv)) %>% 
  summarise(across(c(improv_eob, improv_final, perc), median))
```

Figure \@ref(fig:meanerr) shows the biased (instead of absolute)
estimated deviation, highlighting that the uncertainty range associated
with the conventional method is significantly larger compared to even
the randomized method that only takes 36 weeks, according to subplot (c)
in Figure \@ref(fig:abstract).

```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=8, fig.width=12}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Estimated error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_middle, 
          ncol = 1, nrow = 2,
          labels = c("a)", "b)")) +
  plot_annotation(title = "Estimated error in savings estimation by different M&V methods")

percentiles <- df_TW_A %>%
  filter(method == "conv_diff") %>% 
  summarise(
    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
  )
```

Additionally in the literature, we found another M&V study using over
500+ commercial buildings (on a different open-source dataset), but
their study scope was to compare the predictive performance of a variety
of M&V baseline models from simple weekly mean to more complex machine
learning methods using the normalized mean bias error (NMBE)
[@granderson_accuracy_2016]. Since we also calculated the normalized
savings estimation error, we compared our normalized accuracy for 600+
buildings with the TOWT model and TMY weather againts their TOWT model
prediction assessment in the following table, highlighting that the
model and overall building datasets are performing similarly.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

: Table 2. Savings accuracy comparison with a similar literature

## Non-routine events impact

As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
the selected buildings are affected by operational drift, which is the
reason why we observe baseline electricity usage changes throughout the
two-year period. To evaluate the robustness of the two M&V methods to
the effect of operational drift and non-routine events, we repeated the
M&V run on the original dataset prior to the application of the reset
control. In other words, there *is no intervention applied here*, so the
most reliable M&V method will be the one that detects the closest to 0
kW savings using the measurements. Therefore, the $Acc$ here are
calculated as:

$$
Acc = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$

Figure \@ref(fig:nosaving) shows that the randomized M&V method (both at
the time of satisfying all stopping criteria and at the end of a
two-year period) produced consistent results. We included the absolute
deviation ($|ACC|$) figure in the supplementary material. Similarly to
Figure \@ref(fig:abstract), stopping when meeting all stopping criteria
generally yields slightly larger uncertainty than continuing for 24
months. However, it still yields a more accurate result than the
conventional method. Interestingly, similar to
[@granderson_accuracy_2016] where they obtained a median normalized
deviation around -1.25%, our analysis on the conventional method shows
that the typical deviation is -1.5%, meaning that the typical building
is decreasing energy use slightly in this dataset in the absence of any
known intervention. Thus, if a building analyst were to apply an
intervention to this building and use conventional M&V to measure the
effect, they would typically 'detect' 1.5% energy savings even when the
intervention had no actual impact on energy performance.

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=5, fig.width=12}

#### MEAN-NULL ####
rand_eob_S <- df_seq_fs_null_stable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_null_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_null_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_null_variable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_null_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_null_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                                "conv_diff" = "Conventional", 
                                "rand_eob_diff" = "Randomized", 
                                "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 6), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Estimated effect in fractional savings",
       title = "Estimated effect of non-routine events when using different M&V methods", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-18, 18)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

# Discussion

## TOWT modeling accuracy

To evaluate modeling accuracy, we used the Coefficient of Variation of
Root-Mean Squared Error (CV(RMSE)) as the primary error metric. CV(RMSE)
is a normalized measure, thus it enables direct comparison across
different modeling results. Since we have the ground-true baseline
measurements in the post-retrofit period, we can assess the error caused
by model adjustment (i.e., how accurate is the 'counterfactual' baseline
provided by the two M&V methods). Figure \@ref(fig:towtacc) shows the
difference between the true post-retrofit baseline and adjusted baseline
through TOWT modeling and since the quantity of the training set is the
same, this highlights the impact of the data sampling technique. For
example, if a non-routine event happens (such as a tenant occupying an
entire floor of a building moving out), the TOWT model can yield
significant deviations. This is particularly the case for the
conventional M&V method since it samples continuously throughout the
pre-retrofit period. However, randomized sampling is less impacted by
these events since it samples only 50% of the pre-retrofit baseline
throughout a blocking period. Consequently, the figure shows, there is a
noticeable improvement in modeling accuracy when using the randomized
method.

```{r towtacc, fig.cap="TOWT modeling accuracy distribution for all buildings included (with each data point representing one building)", fig.width=8, fig.height=4}

cvrmse <- df_model_acc_stable %>% 
  bind_rows(df_model_acc_variable) %>% 
  pivot_longer(c(conv, rand), 
               names_to = "method", 
               values_to = "cvrmse") %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                              "rand" = "Randomized\n(24 months)", 
                              "conv" = "Conventional"))

result <- t.test(filter(cvrmse, method == "Randomized\n(24 months)")$cvrmse, 
                 filter(cvrmse, method == "Conventional")$cvrmse, 
                 exact = TRUE,
                 alternative = "less",
                 mu = 0,
                 conf.int = TRUE, 
                 conf.level = 0.90,
                 var.equal = TRUE)$p.value

cvrmse %>% 
  ggplot(aes(x = method, y = cvrmse, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  annotate(geom = "text", 
           x = 1.5, 
           y = 40, 
           size = 4, 
           label = paste0("P value = ", round(result, digits = 3))) +
  geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
  geom_text(data = . %>% 
              group_by(method) %>% 
              summarise(mean = mean(cvrmse, na.rm = T)) %>% 
              ungroup(), 
            aes(x = method, y = mean, label = paste0(round(mean, digits = 0), " %"))) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %")) +
  coord_cartesian(ylim = c(0, 50)) +
  labs(x = NULL, 
       y = NULL, 
       color = NULL, 
       title = "TOWT model fitting accuracy by different M&V methods", 
       subtitle = "CV(RMSE)") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Sampling interval impact

In some cases, particularly those involving large thermal mass and
active thermal energy storage systems or operating continuously, there
can be a substantial time-lagged effect associated with interventions.
This means that the effect of one control strategy may persist, or
'carryover', after switching to another. To mitigate such carryover
effects, building analysts may choose to drop non-consecutive days in
the switchback experiment. For example, consider a heat pump water
heater where the tank is charged at very different times of the day by
the intervention strategy compared to the baseline strategy. When
switching strategy, the tank may be much warmer or cooler than typical
for that strategy, thus influencing subsequent measurements,
particularly the performance measured on the first day (i.e., a
non-consecutive day). In such cases, increasing the sampling interval
(e.g., sample strategies every 2 or 3 days instead of daily) and
discarding measurements from those non-consecutive days can eliminate
carryover. However, this approach reduces the total data collected: for
instance, sampling every 3 days and dropping days with non-consecutive
control strategies would drop 1/6 of measurements on average. Sampling
every 2 days means dropping a quarter of the measurements. Therefore, by
increasing the sampling interval, there is an accuracy penalty due to a
reduction in randomization (how often samples occur), and when
decreasing the sampling interval, there is an accuracy penalty from
dropping more non-consecutive days. The dataset contains no specific
information about the buildings' thermal mass condition, but to quantify
such trade-off, we compared the accuracy of different sampling intervals
and the consequence of dropping the non-consecutive days at the time
when randomized M&V produces a result (after satisfying all stopping
criteria) in Figure \@ref(fig:absinterval).

```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=9, fig.width=12}

#### ABS-INT ####
# sprt sequence plot for different sampling intervals
df_seq_interval_fs_keep_stable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_seq_fs_stable %>% 
              filter(seq == "eob") %>% 
              select(-seq) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "fs"))

df_seq_interval_fs_keep_variable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_seq_fs_variable %>% 
              filter(seq == "eob") %>% 
              select(-seq) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "fs"))

df_drop <- df_seq_interval_fs_drop_stable %>% 
  bind_rows(df_seq_interval_fs_drop_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "drop")

df_keep <- df_seq_interval_fs_keep_stable %>% 
  bind_rows(df_seq_interval_fs_keep_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "keep")

df_MW_A <- df_drop %>% 
  bind_rows(df_keep) %>% 
  left_join(rbind(df_fs_stable %>% 
                    filter(scenario == "ref" & method == "true"), 
                  df_fs_variable %>% 
                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(diff = abs(savings - eob)) %>% 
  filter(is.finite(diff))

p_top <- df_MW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_jitter(alpha = 0.8, 
              size = 0.5, 
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, inf.rm = T)) %>% ungroup(), 
            aes(x = interval, y = median, group = group,
                label = paste0(round(median, digits = 1), " %")), 
            position = position_dodge(width = 0.75), 
            color = "white", 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# with tmy version
df_seq_interval_nm_keep_stable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_sprt_all_stable %>% 
              filter(seq == "eob") %>% 
              select(-c(seq, n_weeks)) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "annual"))

df_seq_interval_nm_keep_variable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_sprt_all_variable %>% 
              filter(seq == "eob") %>% 
              select(-c(seq, n_weeks)) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "annual"))

df_drop <- df_seq_interval_nm_drop_stable %>% 
  bind_rows(df_seq_interval_nm_drop_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "drop")

df_keep <- df_seq_interval_nm_keep_stable %>% 
  bind_rows(df_seq_interval_nm_keep_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "keep")

df_TW_A <- df_drop %>% 
  bind_rows(df_keep) %>% 
  left_join(rbind(df_fs_stable %>% 
                    filter(scenario == "ref" & method == "true"), 
                  df_fs_variable %>% 
                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(diff = abs(savings - eob)) %>% 
  filter(is.finite(diff))

p_bottom <- df_TW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_jitter(alpha = 0.8, 
              size = 0.5, 
              position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
            aes(x = interval, y = median, group = group,
                label = paste0(round(median, digits = 1), " %")), 
            position = position_dodge(width = 0.75), 
            color = "white", 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_bottom, 
          ncol = 1, nrow = 2, 
          align = "hv", 
          labels = c("a)", "b)"),
          common.legend = T, 
          legend = "bottom") +
  plot_annotation(title = "Absolute error in savings estimation of randomized M&V after satisfying all stopping criteria")
```

To quantify the impact of reduced randomness, all measurements are kept
when calculating the distribution shown in the darker-colored set. The
results suggest that without dropping non-consecutive days in subplot
(a), a gradual increase in error from 1.5% to 2.3% (1.2% to 1.9% with
weather normalization in subplot (b) when the sampling interval
increases from daily to weekly intervals. This is due to the fact that
increasing the sample interval means reducing randomization. To quantify
the combined impact of data loss due to preventing the carryover effect,
we generated the lighter-colored set, which drops all non-consecutive
days. As a result, we noticed an increase in estimation error when
sampling daily to 2.4% (to 1.9% when TOWT is fitted), which led to a
similar accuracy compared to weekly intervals. Therefore, considering
this trade-off, we recommend using a two-day sampling interval (or
three-day sampling interval if the building owner prefers less frequent
switchover) if the carryover effect is likely but expected to last less
than one day, and then normalizing the estimation via energy modeling.
The supplementary material contains similar figures showing biased
deviation distribution, as well as the same results when continuing the
randomized M&V for 24 months. Those results indicate the same
conclusion; that sampling every two days and dropping non-consecutive
samples yields the optimal result when carryover effects are present.

Sampling at different intervals and excluding non-consecutive days also
impact other stopping criteria, such as the time needed to satisfy
statistical uncertainty and weather conditions thresholds. In Figure
\@ref(fig:intervaltime), we compared the percentage of buildings
satisfying all stopping criteria under different sampling intervals,
with or without dropping non-consecutive days. The results indicate that
randomized M&V concludes more quickly when all data are retained,
particularly at shorter intervals, as expected. For instance, when
sampling at a one-day interval, 90% of buildings meet all criteria
within 24 weeks if no data are dropped, but this number decreases
significantly to 52% when non-consecutive days (account for 1/3 of the
data) are excluded, and 3% of the buildings require another 12 weeks to
satisfy the weather criterion. As the sampling interval increases, this
difference gradually diminishes: when sampling at a weekly interval,
removing non-consecutive days has no impact on the timeline of meeting
stopping criteria.

```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=8, fig.width=12}
df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
  mutate(group = "drop", 
         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
  select(name, seq, n_weeks, interval, group)

interval_1_keep <- bind_rows(df_sprt_all_stable, df_sprt_all_variable) %>%
  select(c(name, seq, n_weeks)) %>%
  filter(seq != "final") %>%
  mutate(seq = as.factor(seq)) %>%
  mutate(interval = 1, 
         group = "keep")

df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_variable) %>%
  mutate(group = "keep", 
         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
  select(name, seq, n_weeks, interval, group) %>% 
  bind_rows(interval_1_keep)

df_MW_A <- bind_rows(df_keep, df_drop) %>% 
  filter(seq != "final") %>% 
  filter(seq == "eob") %>% 
  mutate(seq = as.factor(seq), 
         seq = recode_factor(seq, 
                             "eob" = "Buildings finishing all criteria"), 
         interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept"), 
         n_weeks = as.factor(n_weeks), 
         n_weeks = recode_factor(n_weeks, 
                                 "12" = "12\nweeks", 
                                 "24" = "24\nweeks", 
                                 "36" = "36\nweeks", 
                                 "48" = "48\nweeks")) %>% 
  group_by(n_weeks, group, interval) %>% 
  summarise(n = n()) %>%
  ungroup()

levA <- levels(df_MW_A$n_weeks)
levB <- levels(df_MW_A$group)
levC <- levels(df_MW_A$interval)

# 2. Expand to all combinations of factorA and factorB
all_combinations <- expand_grid(
  n_weeks = levA,
  group = levB, 
  interval = levC, 
)

df_MW_A <- all_combinations %>%
  left_join(df_MW_A, by = c("n_weeks", "group", "interval")) %>% 
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  group_by(group, interval) %>% 
  mutate(n = ifelse(n > sum(n) * 0.01, n, 0), 
         perc = n / sum(n) * 100) %>% 
  ungroup()

df_MW_A %>% 
  ggplot(aes(x = n_weeks, y = perc, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~interval, nrow = 2) +
  geom_text(aes(x = n_weeks, y = perc + 2, label = paste0(round(perc, digits = 0), "%"), group = group), 
            position = position_dodge(1), 
            size = 4) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Percentage of buildings satisfying all stopping critiera", 
       subtitle = "Counted at the end of each 12-week blocking period") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Sampling ratio impact

In our previous study [@raftery2024], we recommended that building
owners adjust the sampling ratio after meeting all stopping criteria, as
continuing baseline sampling helps maintain M&V accuracy with limited
impact on the energy savings actually achieved by the building. To
further examine its impact on M&V accuracy, Figure \@ref(fig:abscont)
compares two sampling ratios (20%/80% and 10%/90%) and two associated
scenarios with the previously shown results when sampling at 50%/50%.
The "Continued" group (shown in a lighter color) follows the standard
randomized M&V framework with an initial 50%/50% sampling ratio, then
switches to the indicated sampling ratio after satisfying the stopping
criteria, continuing this until the end of the two-year period. In
contrast, the "Randomized" group uses the indicated ratio from the
outset and continues throughout the two year. Subplot (a) shows that
absolute error is significantly lower without the initial 50%/50%
sampling, due to baseline drift and non-routine events that most
buildings experience over two years (as noted in Section
\@ref(filter-and-clean-dataset)). Subplot (b) confirms this finding even
after weather normalization.

```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=8, fig.width=12}

#### ABS-CONT ####
df_MW_S <- df_cont_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs), 
         diff_all = abs(savings - cont_fs_all)) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_stable %>% 
              filter(method != "conv") %>% 
              pivot_wider(names_from = method, values_from = savings) %>% 
              mutate(values = abs(true - rand), 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for the variable subset
df_MW_V <- df_cont_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs), 
         diff_all = abs(savings - cont_fs_all)) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_variable %>% 
              filter(method != "conv") %>% 
              pivot_wider(names_from = method, values_from = savings) %>% 
              mutate(values = abs(true - rand), 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for combined dataset
df_MW_A <- rbind(df_MW_S, df_MW_V)

p_top <- df_MW_A %>% 
  mutate(ratio = as.factor(ratio), 
         ratio = recode_factor(ratio, 
                               "0" = "Randomized\n(50/50)", 
                               "1_1" = "Continued\n(20/80)", 
                               "1_2" = "Randomized\n(20/80)", 
                               "2_1" = "Continued\n(10/90)", 
                               "2_2" = "Randomized\n(10/90)")) %>% 
  ggplot(aes(x = ratio, y = values, fill = ratio)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(ratio) %>% 
              summarise(median = median(values)) %>% 
              ungroup(), 
            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(), 
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
df_TW_S <- df_cont_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_tmy), 
         diff_all = abs(savings - cont_tmy_all)) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_tmy_stable %>% 
              left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
              mutate(values = abs(savings - rand), 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for the variable subset
df_TW_V <- df_cont_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_tmy), 
         diff_all = abs(savings - cont_tmy_all)) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_tmy_variable %>% 
              left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
              mutate(values = abs(savings - rand), 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for combined dataset
df_TW_A <- rbind(df_TW_S, df_TW_V)

p_bottom <- df_TW_A %>% 
  mutate(ratio = as.factor(ratio), 
         ratio = recode_factor(ratio, 
                               "0" = "Randomized\n(50/50)", 
                               "1_1" = "Continued\n(20/80)", 
                               "1_2" = "Randomized\n(20/80)", 
                               "2_1" = "Continued\n(10/90)", 
                               "2_2" = "Randomized\n(10/90)")) %>% 
  ggplot(aes(x = ratio, y = values, fill = ratio)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75) +
  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(ratio) %>% 
              summarise(median = median(values)) %>% 
              ungroup(), 
            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_bottom, 
          ncol = 1, nrow = 2,
          labels = c("a)", "b)")) +
  plot_annotation(title = "Absolute error in savings estimation by different sampling ratio")
```

We further illustrated this point in the supplementary material where we
plotted results using only the 'stable' building subset (66 buildings),
as defined in Section \@ref(filter-and-clean-dataset). Since they are
less influenced by non-routine events, the differences between sampling
strategies are much reduced in this subset and can be largely corrected
through weather normalization modeling.

As shown in the previous Figure \@ref(fig:abstract), the highest
accuracy is achieved when sampling at a 50%/50% ratio. Therefore, we
continue to recommend that building owners follow the proposed framework
and maintain a 50%/50% sampling ratio until all stopping criteria are
met. The additional results from Figure \@ref(fig:abscont) further add
to the guidance: if building baseline changes, retrofit savings
estimation should simply discard the existing baseline measurements and
revert back to 50%/50% sampling ratio as described in our framework. On
the other hand, if the building owner has decided to test for two years
and would like to capitalize more savings at the same time, sampling
between the baseline and intervention at 10%/90% from start is
recommended.

## Blocking period impact

Selecting different blocking periods affects how frequently the M&V
analyst evaluates stopping criteria and may influence the decision when
a savings is detected at a sufficient level of certainty. Here, we
summarize the advantages and disadvantages of shorter versus longer
blocking periods. As discussed in our previous study, the second step of
this randomized M&V method involves choosing an appropriate blocking
period to minimize potential biases caused by incomplete randomization.
For most practical scenarios, we recommend a blocking period of
approximately 12 to 16 weeks, as it typically covers the duration of a
season, ensuring a representative sampling of operational conditions. In
general, longer blocking periods are advantageous when evaluating
multiple intervention strategies or using sampling intervals greater
than one day. In such cases, extended blocking ensures a balanced
distribution of both day-of-week and outdoor weather conditions.

Opting for a shorter blocking period, such as 6 weeks, has the potential
benefit of allowing the stopping criteria to be met earlier. We plotted
the percentage of buildings finishing all stopping criteria when the
blocking period is set as 6 weeks in Figure \@ref(fig:6weekblock). As a
result, when compared to Figure \@ref(fig:intervaltime), the only
difference we found is when sampling at a daily interval and not
dropping non-consecutive days, 12% of buildings can get an estimate
earlier. This means those buildings are likely to meet all stopping
criteria between 12 weeks and 18 weeks but when sampled over a 12-week
block, the savings can only be determined at week 24. Although sampled
baseline and intervention days can be different when the blocking period
is changed, the results indicate most buildings still require 9 months
to meet outdoor weather conditions when starting M&V from January.

```{r 6weekblock, fig.cap = "Time required to satisfy all stopping criteria when using a 6-week blocking period"}

knitr::include_graphics(paste0(fig_path, "6week_time.png"))
```

Sampling at a shorter blocking period also means that the analyst must
perform the hypothesis tests more frequently, which can necessitate
adjusting the significance level using methods such as the Bonferroni
correction to control for increased type-I error risk. On the plus side,
a shorter blocking period can reduce the adverse impact of missing or
problematic data. For example, if an intervention requires fine-tuning
after initial implementation, removing data from the first blocking
period would be less detrimental compared to longer blocking periods,
preserving the overall integrity of the experiment.

## M&V starting time impact

In this study, we set the baseline period to begin on January 1st, which
offers the advantage of capturing colder outdoor weather conditions
early in the M&V process. However, when the baseline measurements miss
the initial winter period, meeting the required weather condition
criteria can potentially take a longer time. A previous study also
showed that M&V accuracy, especially model fitting, depends on both
monitoring duration and starting month, with optimal modeling achieved
when sufficient weather variation is captured [@singh_predicting_2014].
To assess this impact in our switchback randomization, we repeated the
randomized M&V analysis for all buildings using a new start date in
March. As shown in Figure \@ref(fig:startT), the new M&V process results
in a noticeable delay in satisfying the stopping criteria. Compared to
Figure \@ref(fig:intervaltime), where a January start allowed most
buildings to finish within 36 weeks, a March start extended the required
duration to 48 weeks for the majority of buildings. Furthermore, when a
longer sampling interval is used or non-consecutive days are excluded
(to avoid carryover effects), some buildings may require up to 60 weeks
to complete the randomized M&V. As expected, to collect sufficient data
representing winter conditions, the sampling may need to continue until
the next January (around week 48). However, as the overall data
collection period extends, the influence of dropping non-consecutive
days diminishes and becomes less of a concern.

```{r startT, fig.cap = "Time required to satisfy all stopping criteria starting from March"}

knitr::include_graphics(paste0(fig_path, "startT.png"))
```

Similarly, we evaluated the time required to satisfy all stopping
criteria when using a shorter blocking period, and included the plot in
the supplementary material. We found that compared to Figure
\@ref(fig:startT), a significant proportion of buildings completed the
randomized M&V by week 42 (mid-December), which is one blocking period
earlier than having a 12-week blocking period. This result further
highlights that, in general, starting M&V in the middle of the winter or
summer periods will yield results spanning the full range of weather
conditions faster than starting at other times of year, such as during a
swing season.

## Limitations

Aside from the general limitation that this method only applies to the
subset of interventions in buildings that are easily switchable, we
identify two key limitations in this study:

1.  The simulated intervention (chilled water temperature reset) remains
    somewhat generic and simplistic, considering the diverse building
    types and climate zones in the BDG2 dataset. As the primary focus is
    to accurately detect the intervention (or any intervention) effect
    so for simplicity, we applied the parameters listed in Table 1
    uniformly across all buildings.
2.  For the main analysis, we assumed no carryover effect and that a
    daily sampling interval would suffice for most commercial buildings,
    but certainly exceptions exist. However, in the discussion, we
    addressed such concern by comparing different sampling strategies
    and the impact of dropping measurements from non-consecutive days.

# Conclusion

This research demonstrated the application of a randomized
whole-building measurement and verification (M&V) method, comparing its
performance to the conventional approach described in the IPMVP and in
ASHRAE Guideline 14, using a large, open-source commercial building
dataset. The randomized M&V method leverages the randomized experimental
design concept from other scientific fields, along with statistical
sequential inference techniques, to determine when savings are detected
at sufficient certainty. We used a virtual control retrofit
case---resetting the chilled water setpoint based on outdoor weather
conditions---and applied it to over 600 filtered commercial buildings.
By comparing the savings estimations of the conventional method with the
novel randomized method, we found that the randomized approach provides
faster and more robust savings estimations.

Specifically, we showed that throughout 11 different locations assessed
in this study, the randomized M&V can provide a savings estimation after
36 weeks (with the majority finished by 24 weeks) once all stopping
criteria are satisfied. In contrast, the conventional method requires a
full range of baseline and intervention measurements under normal
operating conditions, typically taking 6 - 9 months for each phase. Most
importantly, we verified that with a reduced M&V timeline, the
randomized method can estimate savings more accurately by showing that
the absolute error is only 1 - 2% for a typical building. Whereas for
the much longer two-year conventional method, the estimation error was
approximately 5% in a typical building. We also evaluated the impact of
non-routine events, especially more gradual and subtle operational drift
on the proposed M&V method and the results show that baseline changes in
the post-retrofit period can deviate savings estimated using the
conventional method. In contrast, we found that those events have a very
negligible impact on the savings estimated using the randomized method,
demonstrating robustness to this issue. Furthermore, we consider
scenarios where there is a known carryover effect from switching between
strategies in the building and assess the impact of dropping samples
when the strategy is non-consecutive and varying the sampling interval.
Through that process, we showed that for the typical building, if
carryover is present, using a 2-day sampling interval and dropping days
with non-consecutive control strategies operating yields the optimal
design for the median building in this dataset. Lastly, we investigated
the time required to complete the randomized M&V method under different
scenarios. We found that measuring a sufficient range of outdoor weather
conditions remains the most stringent criterion. However, even in the
most non-ideal case, the randomized method can still produce an estimate
much faster than two years, and produces an estimate with far lower
error.

# CRediT authorship contribution statement

**Aoyu Zou**: Conceptualization, Data curation, Formal analysis,
Methodology, Investigation, Software, Writing - original draft,
Writing - review & editing. **Paul Raftery**: Conceptualization, Funding
acquisition, Formal analysis, Methodology, Investigation, Project
administration, Supervision, Writing - review & editing. **Stefano
Schiavon**: Conceptualization, Methodology, Investigation, Supervision,
Writing - review & editing. **Carlos Duarte**: Methodology,
Investigation, Supervision, Writing - review & editing. **Gail Brager**:
Supervision, Writing - review & editing.

# Reproducibility

A reproducible example with analysis code is available (MIT license) at
<https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.

# Declaration of competing interest

The authors declare that they have no known competing financial
interests or personal relationships that could have appeared to
influence the work reported in this paper.

# Acknowledgements

# References