### 2025-07-15 code/manuscript.Rmd
updaded review for manuscript and updated readme file

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou:
      institute: cbe
      correspondence: true
  - name: Paul Raftery
    institute: cbe
  - name: Stefano Schiavon
    institute: [cbe, arch, cee]
  - name: Carlos Duarte
    institute: cbe
  - name: Gail Brager
    institute: [cbe, arch]
institute:
  - cbe: Center for the Built Environment, University of California Berkeley, USA
  - arch: Department of Architecture, College of Environmental Design, University of California, Berkeley, USA
  - cee: Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA
abstract: Conventional measurement and verification (M&V) methods for estimating energy savings rely on comparing pre- and post-retrofit performance. They are often time-consuming and unreliable, especially when non-routine events, such as step changes or more gradual changes in building operation, occur during the M&V process. When those events are unrelated to the retrofit intervention but may significantly affect building energy consumption, then when the analyst applies the conventional M&V method the results can be confounded. In this study, we demonstrated that switchable interventions, such as most HVAC control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that, compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction). Additionally, we found that when non-routine events are present (e.g., occupancy pattern change), the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
    number_sections: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'

scholarly-metadata:
  affiliations-location: footnote

editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../paper"
    )
  })
---

```{r setup, include = FALSE, cache = FALSE}

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      # cache = T,
                      dpi = 300,
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/manuscript/figs/")

# str(knitr::opts_chunk$get()) # see for all options

```

```{r prep, include = FALSE, cache = FALSE}
#### LIBRARIES ####
require(pacman)

# load packages using pacman
pacman::p_load(tidyverse, lubridate, here, stats, zoo, scales, lvplot, ggpubr, gridExtra, patchwork, RColorBrewer)

# turn off scientific notation
options(scipen = 999, digits = 15)

# set directory
here::i_am("manuscript.rmd")

# set default theme for ggplot
theme_set(theme_minimal())

# define base ggplot theme
theme_update(plot.title = element_text(size = 16, colour = "grey20", face = "bold", hjust = 0.5),
             plot.subtitle = element_text(size = 14, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
             plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
             plot.background = element_rect(fill = "white", colour = NA),
             panel.grid.minor = element_blank(),
             panel.grid.major = element_blank(),
             axis.text = element_text(size = 12),
             legend.text = element_text(size = 12),
             strip.text = element_text(size = 12, color = "grey20", face = "bold"),
             strip.background = element_blank())

# colors
ls_colors <- c("Baseline" = "grey10",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional" = "grey70", 
               "Randomized" = "#99d8c9", 
               "Randomized\n(24 months)" = "#6baed6",
               "Randomized\n(50/50)" = "#6baed6",
               "Randomized\n(20/80)" = "#6baed6", 
               "Randomized\n(10/90)" = "#6baed6", 
               "Continued\n(20/80)" = "#c6dbef", 
               "Continued\n(10/90)" = "#c6dbef",
               "Daily\nsampling" = "#66c2a4",
               "2-day\nsampling" = "#66c2a4",
               "3-day\nsampling" = "#66c2a4",
               "7-day\nsampling" = "#66c2a4",
               "Dropped" = "#41ae76",
               "Kept" = "#006d2c",
               "Dropped (24 months)" = "#4292c6",
               "Kept (24 months)" = "#08519c",
               "Buildings finishing randomized M&V" =  "#99d8c9",
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

# parameters
ctr_params <- list(peak_hours = 10:16,
                   chwl_perc = 0.25,
                   step_perc = 0.08,
                   conv_swt = 6,
                   weather_knots = c(15, 25),
                   swt_knots = c(12, 7),
                   coe_peak = 0.8,
                   coe_off = 1.2,
                   enable_temp = 8)

occ_params <- list(change_start = 5,
                   change_end = 8,
                   change = 20)

#### FUNCTIONS ####
# Function defined to read downloaded tmy files
get_tmy <- function(all_sites, readfile_path){

  df_tmy <- data.frame()

  for (site in all_sites){
    df <- read_csv(paste0(readfile_path, "tmy/", str_glue("{site}.epw")),
                   skip = 8, col_types = "ddddd-d---------------------------------",
                   col_names = c("year", "month", "day", "hour", "min", "tmy")) %>%
      mutate(year = 2017,
             time = ymd_h(paste(paste(year, month, day, sep = "-"), hour, sep = " ")),
             temp = tmy) %>%
      dplyr::select(time, temp) %>% 
      mutate(site = site)

    df_tmy <- rbind(df_tmy, df)
  }

  return(df_tmy)
}

# Function defined to add chwst reset intervention
run_reset <- function(df_baseline){

  mean <- mean(df_baseline$base_eload, na.rm = T) * ctr_params$chwl_perc

  grad <- (ctr_params$swt_knots[2] - ctr_params$swt_knots[1]) / 
    (ctr_params$weather_knots[2] - ctr_params$weather_knots[1])

  interc <- ctr_params$swt_knots[2] - (ctr_params$weather_knots[2] * grad)

  df_interv <- df_baseline %>% 
    mutate(swt = t_out * grad + interc, 
           chwl = mean,
           hour = hour(datetime)) %>% 
    mutate(swt = ifelse(swt > ctr_params$swt_knots[1], ctr_params$swt_knots[1], ifelse(swt < ctr_params$swt_knots[2], ctr_params$swt_knots[2], swt)), 
           temp_savings = ifelse(t_out >= ctr_params$enable_temp, (swt - ctr_params$conv_swt) * ctr_params$step_perc, 0), 
           time_adj = ifelse(hour %in% ctr_params$peak_hours, ctr_params$coe_peak, ctr_params$coe_off), 
           perc_savings = temp_savings * time_adj, 
           savings = chwl * perc_savings, 
           interv_eload = base_eload - savings) %>% 
    select(datetime, base_eload, interv_eload, t_out)

  return(df_interv)
}

# Function defined to interpolate NAs
run_interpo <- function(df_all){

  na_counts <- df_all %>%
    mutate(date = date(timestamp)) %>%
    group_by(date) %>%
    summarize(na_hours = sum(is.na(eload)))

  # Filter out days with more than half of the hours having NAs
  valid_days <- na_counts %>%
    filter(na_hours <= 12) %>%
    pull(date)

  df_filtered <- df_all %>%
    filter(date(timestamp) %in% valid_days)

  df_filtered <- df_filtered %>%
    mutate(across(c(eload, t_out), ~ zoo::na.approx(., na.rm = FALSE))) %>% 
    rename(datetime = timestamp, 
           base_eload = eload)

  return(df_filtered)
}

# Function defined to adjust the plot scale
get_scale <- function(eload, range = 2){

  min_y <- mean(eload, na.rm = T) - range * sd(eload, na.rm = T)
  max_y <- mean(eload, na.rm = T) + range * sd(eload, na.rm = T)

  return(c(min_y, max_y))
}
```

```{r readdata, include=FALSE}

#### READ DATA ####
df_loc <- read.csv("../readfiles/loc_map.csv")

# stable dataset
readfile_stable <- str_glue("../readfiles/stable/")
fig_path = "../figs/manuscript/"

df_energy_stable <- read_rds(paste0(readfile_stable, "df_energy.rds"))
df_weather_stable <- read_rds(paste0(readfile_stable, "df_weather.rds"))
df_sprt_all_stable <- read_rds(paste0(readfile_stable, "df_sprt_all.rds"))
df_seq_fs_stable <- read_rds(paste0(readfile_stable, "df_seq_fs.rds"))
df_fs_stable <- read_rds(paste0(readfile_stable, "df_fs.rds"))
df_cont_stable <- read_rds(paste0(readfile_stable, "df_cont.rds"))
df_fs_null_stable <- read_rds(paste0(readfile_stable, "df_fs_null.rds"))
df_seq_fs_null_stable <- read_rds(paste0(readfile_stable, "df_seq_fs_null.rds"))
df_model_acc_stable <- read_rds(paste0(readfile_stable, "df_model_acc.rds"))
df_fs_tmy_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy.rds"))
df_fs_tmy_null_stable <- read_rds(paste0(readfile_stable, "df_fs_tmy_null.rds"))
df_interval_drop_stable <- read_rds(paste0(readfile_stable, "df_interval_drop.rds"))
df_interval_keep_stable <- read_rds(paste0(readfile_stable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_stable <- read_rds(paste0(readfile_stable, "df_seq_interval_tl_keep.rds"))

all_sites_stable <- df_energy_stable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_stable <- df_energy_stable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_stable <- df_energy_stable %>%
  select(name) %>%
  distinct(name)

# variable dataset
readfile_variable <- str_glue("../readfiles/variable/")

df_energy_variable <- read_rds(paste0(readfile_variable, "df_energy.rds"))
df_weather_variable <- read_rds(paste0(readfile_variable, "df_weather.rds"))
df_sprt_all_variable <- read_rds(paste0(readfile_variable, "df_sprt_all.rds"))
df_seq_fs_variable <- read_rds(paste0(readfile_variable, "df_seq_fs.rds"))
df_fs_variable <- read_rds(paste0(readfile_variable, "df_fs.rds"))
df_cont_variable <- read_rds(paste0(readfile_variable, "df_cont.rds"))
df_fs_null_variable <- read_rds(paste0(readfile_variable, "df_fs_null.rds"))
df_seq_fs_null_variable <- read_rds(paste0(readfile_variable, "df_seq_fs_null.rds"))
df_model_acc_variable <- read_rds(paste0(readfile_variable, "df_model_acc.rds"))
df_fs_tmy_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy.rds"))
df_fs_tmy_null_variable <- read_rds(paste0(readfile_variable, "df_fs_tmy_null.rds"))
df_interval_drop_variable <- read_rds(paste0(readfile_variable, "df_interval_drop.rds"))
df_interval_keep_variable <- read_rds(paste0(readfile_variable, "df_interval_keep.rds"))
df_seq_interval_fs_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_drop.rds"))
df_seq_interval_fs_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_fs_keep.rds"))
df_seq_interval_nm_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_drop.rds"))
df_seq_interval_nm_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_nm_keep.rds"))
df_seq_interval_tl_drop_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_drop.rds"))
df_seq_interval_tl_keep_variable <- read_rds(paste0(readfile_variable, "df_seq_interval_tl_keep.rds"))

all_sites_variable <- df_energy_variable %>%
  select(site) %>%
  distinct() %>%
  arrange(site)

all_types_variable <- df_energy_variable %>%
  select(type) %>%
  mutate(type = as.factor(type)) %>%
  distinct()

all_names_variable <- df_energy_variable %>%
  select(name) %>%
  distinct(name)

# read functions
function_path <- "../functions/"
source(paste0(function_path, "model_fit.R"))
source(paste0(function_path, "model_pred.R"))
source(paste0(function_path, "prepost_plot.R"))

S_building <- nrow(df_fs_tmy_stable)
V_building <- nrow(df_fs_tmy_variable)
A_building <- S_building + V_building
```

### Graphical abstract {.unnumbered}

```{r graphicalabstract, out.width = "100%"}

knitr::include_graphics(paste0(fig_path, "graphical_abs.jpg"), dpi = 900)
```

### Keywords {.unnumbered}

Measurement and verification (M&V); Controls; Randomized switchback
design; Uncertainty quantification; Building energy analysis; Heating
ventilation and air conditioning

### Highlights {.unnumbered}

-   Randomized switchback method leads to time reduction and increased
    accuracy
-   Addressed the influence of non-routine events and also more subtle
    long-term 'operational drift' in buildings
-   Investigated the trade-off of selecting longer sampling intervals to
    account for carryover effect and weather dependency of the method
-   Discussed the impact of selecting different sampling ratio on the
    M&V accuracy

# Introduction

## Conventional M&V

Measurement and Verification (M&V) is the process of quantifying
savings, most commonly energy savings, but also potentially water and
demand savings, from efficiency projects by comparing actual consumption
against an established baseline, while adjusting for varying factors
such as weather and occupancy. This process ensures that improvements in
energy performance are accurately evaluated. In the United States,
practitioners often refer to ASHRAE Guideline 14, the International
Performance Measurement and Verification Protocol (IPMVP), and the
Federal Energy Management Program (FEMP) for standard guidelines
[@doe_mv_2008; @efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level, where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. In conventional practice, the process
begins with baseline measurements taken during the year before
implementing the energy-efficiency retrofit, followed by the same
measurement procedure during the year after the intervention. After
collecting two years of data, an M&V analyst fits an energy prediction
model, using variables such as outdoor temperature and time
[@mathieu_quantifying_2011] to project baseline energy consumption in
the post-retrofit time period (but without the retrofit itself). The
difference between this counter-factual baseline and the measured energy
consumption during the actual post-retrofit (intervention) period is the
energy savings. However, other changes unrelated to the implemented
strategies, such as occupancy patterns, can also contribute to the
energy consumption difference. Although building analysts can apply
routine adjustments to account for expected variations in independent
variables commonly included in energy models, such as weather conditions
and time, it is hard to accurately adjust for building performance
change caused by non-routine events. Those events refer to less
predictable and infrequent events that impact energy consumption but are
unrelated to the specific intervention under study. Specifically, when
using the conventional method, an M&V analyst should account for two
additional types of baseline changes beyond routine adjustments: (1)
known static changes to building operation, which are considered as
typical non-routine events: including renovation, equipment addition or
removal, and changes related to tenant turnover; and (2) gradual
changes, which often occur subtly and unknown due to lack of accurate
measurements to quantify, such as the incremental adoption of LED
lighting or more efficient plug loads. Identifying and quantifying these
changes are particularly labor-intensive and challenging. Consequently,
those limitations reduce the feasibility of M&V and complicate the
quantification of estimation uncertainty, thus impacting the accuracy
and timeliness of savings assessments.

## Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston_comparison_2019; @duan2013;
@gupta2019; @mie2022; @munro; @raseduzzaman2017]. An early application
of this randomized design to the field of M&V in buildings was first
used in Raftery et al. [@raftery2018], which compared the energy
performance of two different supply air temperature reset strategies in
a building. Through random sampling, the study reached a conclusion of
29% energy cost savings after just 6 months. Subsequent improvements to
the M&V method included adding a sequential evaluation framework and
defining stopping criteria to end the M&V period early after reaching a
target level of uncertainty, using blocked randomization by weekday. The
full framework is described in a previously published study
[@raftery2024] along with a case study application to a real building,
and other example use cases. In summary, this method provides M&V
analysts with a randomized schedule that alternates between baseline and
intervention implementation while ensuring balanced sampling across days
of the week and seasons. For example, given a 10-week M&V period for one
intervention, the balanced randomized schedule would equally sample five
Mondays with the baseline strategy in operation and five Mondays with
the intervention strategy in operation.

The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off (e.g., HVAC control retrofits), which is a small subset of
all the interventions performed in buildings. However, for all
applicable use cases, it allows analysts to detect energy savings
sequentially shortly after the test begins. In addition, once the
desired estimate uncertainty target is achieved, analysts can terminate
the M&V study and switch to 100% intervention. The key advantage of
randomization is that when control strategies are sampled with equal
probability, the influence of confounding factors (e.g., occupancy
change---a common non-routine event---and other more subtle long term
changes in building energy use) are likely to be evenly distributed
among measurements, leading to a more accurate and unbiased assessment
of the intervention effect.

## BDG2 dataset

To demonstrate the differences between the two M&V methods, we used the
Building Genome Dataset 2 (BGD2), which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies, acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and real operational measurements
between 2016 and 2017 from across North America and Europe, making it
one of the most comprehensive collections of building-related data
available for scientific use. The dataset includes various commercial
building types such as offices, education facilities, public, and retail
buildings, and provides detailed information on their physical
characteristics (e.g. energy ratings, heating types and floor area) and
hourly measurements of chilled and hot water, electricity, gas usage as
well as site outdoor weather conditions. In this study, we mostly used
the metered electricity and outdoor weather measurements for evaluating
the M&V methods.

## Literature review

### Whole building approach

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required.
Moreover, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have a Coefficient of
Variation of the Root Mean Square Error, CV(RMSE), lower than 25% for
12-month data collection. Thus, most research related to M&V for the
whole-building approach focuses on the accuracy of baseline modeling,
exploring model performance from simple regression models to more
complex machine learning techniques. Alrobaie and Krarti
[@alrobaie_review_2022] reviewed various models suitable for M&V
applications as well as selected input features. Gallagher et al.
[@gallagher_development_2018] provided a well-thought-out methodology to
apply machine learning models for M&V use cases. Researchers at LBNL
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies emphasized the
uncertainty associated with the model-fitting process, a key factor in
accurately determining energy savings. Furthermore, other researchers
addressed this issue by leveraging statistical formulation and inference
to improve baseline energy models [@burkhart_measurement_2014;
@heo_calibration_2012; @walter_uncertainty_2014]. However, a gap still
remains in the literature regarding how to rigorously quantify the
uncertainties directly associated with calculated savings. For instance,
this includes accounting for the potential bias that the baseline model
might deteriorate over time, as well as the model no longer accurately
representing the building energy behavior due to changes in operation,
occupancy, equipment performance, etc.

### Changes in building baseline measurements

As mentioned earlier, non-routine events are changes in a building that
influence its energy usage but are unrelated to the intervention being
studied by the M&V. Current guidelines define non-routine events as
'static factors' that need to be adjusted after projecting the baseline
into the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a
reminder that non-routine events should be reported and properly
adjusted. For example, IPMVP requires the facility owner and the M&V
analyst to periodically perform inspections of all equipment and
operations during the reporting period, which is labor-intensive and
error-prone. ASHRAE Guideline 14 recommends performing engineering
calculations or computer software simulations to adjust the
post-retrofit baseline. Additionally, it is relatively rare to have
access to all measurements needed for adjustments, such as occupancy,
and thus the analyst normally assumes those factors remain unchanged
throughout the study. For example, some studies use linear interpolation
to estimate counter-factual baseline for the demand response program
[@beil_round-trip_2015; @keskar_commercial_2020]. Consequently, it is
inaccurate to assume no change in the operating conditions during the
response period [@huang_experimental_2023]. This is a limitation in a
field study, and it is important to require matched groups to control
for exogenous factors beyond weather (which is usually taken into
account) when comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mainly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. We found a similar concept briefly defined in
Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during a
multi-year M&V project. However, currently there are no detailed
instructions provided on how to account for them. If M&V analysts are
unaware of such changes and lack an appropriate adjustment method (e.g.,
replacing filters before the intervention begins), the savings could be
overestimated as the decreased energy use is incorrectly attributed to
the intervention rather than the filter replacement. An operational
drift has not been formally defined in the literature or standards; in
this study, we consider it as a special type of non-routine event, as it
also represents a change in facility operation that occurs gradually
over a longer period of time.

## Objectives

One goal of an M&V project is to determine the effect, typically energy
savings, of an intervention in the building. In this study, we focus on
switchable interventions. An example of such an intervention is a
control strategy that strategically adjusts the chilled water plant's
supply water temperature with proven ability to save energy
[@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. We used a simple
version of the control strategy in which the setpoint is adjusted based
on outdoor temperature (see description in Section
\@ref(apply-virtual-intervention)). By conducting such an analysis, we
aim to:

1.  Determine how much more accurately the randomized M&V method would
    estimate the savings of the intervention compared to the
    conventional M&V, and how much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset.

3.  Make the implementation of the proposed randomized M&V method
    open-source.

The significance of this study is that the BDG2 dataset contains
real-world building energy usage data collected over two years, which
inherently contains varying degrees of usage changes. In addition, by
virtually introducing intervention effects into the existing dataset, we
can calculate the ground-truth energy savings, which serve as a
reference point for method comparison. Furthermore, we ensured the
reproducibility of the method by making the analysis code open source,
including randomized schedule generation, sequential statistical
analysis, energy modeling and normalized saving calculation. This also
covers the extended discussion below on sensitivity analysis and other
use cases of the randomized method, such as changing sampling ratio and
sampling intervals. Using the available open resources, building
analysts should be able to seamlessly integrate and apply them in their
own M&V projects.

# Method

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section. As a summary, we
utilized the historical measurements from a dataset that contains a
large number of real commercial buildings as the pre-retrofit baseline
and simulated an intervention strategy. Then, we applied both
conventional and randomized M&V methods to the modified dataset to
estimate intervention savings, and compare which one produces results
closer to the simulated intervention effect.

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="60%"}

knitr::include_graphics(paste0(fig_path, "flowchart.png"))
```

## Filter and clean dataset

We extracted the electricity measurements from the BDG2 dataset and
filtered out all qualified buildings based on the following criteria:

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to 1.5 months of missing days.

2.  Mean electricity usage \> 0 kWh: target buildings should have active
    electricity usage.

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage ($E$) in the second year
    ($E_{2017}$) should be less than 25% of that in the first year
    ($E_{2016}$).

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrence_berkeley_national_laboratory_buildings_2015;
    @mathew_big-data_2015].

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water set point reset
    control.

As a result, we kept a total of 639 buildings from 10 different
locations for the subsequent analysis, and we included a more detailed
summary of the building features in the supplementary material. For all
qualified buildings, we also create a feature to identify buildings with
very stable electricity usage overall by identifying those with no
statistically significant difference (P-value \> 0.05) between their
pre- and post-retrofit electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings, corresponding to 11%. While we did
not adjust for weather differences between the two years when generating
this high-level feature, this result highlights that the assumption that
buildings are typically stable over the timeframe involved in
conventional M&V methods is rarely true; in 89% of the cases this
assumption is not valid. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events---especially operational drifts---since the
assessment, even under ideal situations, lasts for two years. According
to the statistics shown here, it is a typical case observed in real
buildings applications.

## Apply **virtual** intervention

The proposed control intervention resets the chiller supply temperature
based on the outdoor temperature, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy adjusts the water supply
temperature dynamically, resetting it from 7°C to 12°C. We included a
visual summary of the algorithm in the supplementary material.

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings:$$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
$E_{CHW}$ further consumes around 50% of the HVAC electricity, or 25% of
the total building electricity [@us2012commercial]. We recognize that
this assumption, used for the scope of this analysis, largely simplifies
the diverse energy usage across various building types. Typically, the
savings from an intervention are not proportional to the building's
hourly electricity usage, which is generally a challenge for M&V. To
address this, we mapped the resulting electricity savings as a
percentage of the plant's normal operation, calculated as its mean
electricity usage over the two-year period. This percentage is
influenced by factors such as outdoor temperature ($OAT$), intervention
supply water temperature ($T_{interv}$), baseline supply water
temperature ($T_{base}$) and hour of the day ($\delta_{occ}$, binary
indicator whether during peak hours from 9 AM to 4 PM).

$$
\text{perc = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below, they were obtained from a trial supply water setpoint reset
experiment on a campus office building [@zou_quantifying_2025]. For
simplicity, those parameters were not rigorously calibrated for each
building and were applied uniformly in the filtered dataset. Interested
readers can also change the parameters in our open-source code to
simulate different scenarios.

| Parameter              | Description                                                      | Value |
|------------------|------------------------------------|------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Table 1. Parameters for calculating the intervention savings.

## M&V methods and results analysis

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post-comparison on a 12-month baseline and 12-month
intervention timeline. The randomized method defines sampling
requirements as (a) use a daily sampling interval with the sampling time
at midnight each day and (b) block by day of the week with a block
period of 12 weeks.

We used the following four stopping criteria: (a) a minimum and maximum
of 12 and 108 weeks, respectively. The randomized schedule covers the
entire two-year period but the stopping criteria enable an early stop at
the end of the satisfied blocking period. (b) At least 80% of the
dry-bulb temperature range in the annual TMY data is sampled by both
strategies. (c) 90% confidence that energy savings exceed or do not
exceed 0% using the sequential probability ratio test (SPRT). Medium
effect size (d = 0.5) quantified by Cohen's d and calculated SPRT
statistics either fall below the lower threshold or exceed the upper
threshold. (d) sample strategies at an equal sampling ratio (50%
baseline, 50% intervention).

To summarize the M&V results and associated error to the methods, we
present the distribution of savings estimation error using boxen plots,
also known as letter plots, which is an advanced variation of the box
plot designed to extend beyond the interquartile range (IQR) by
progressively dividing the data into smaller percentiles, revealing more
detail in the tails of the distribution. We set the division parameter
to k = 4, meaning the entire box area represents the data distribution
from 6.25% to 93.75%. As the steps move closer to the center line (50%
median), the distribution range progressively narrows with the next step
representing 12.5% to 87.5%, followed by 25% to 75%. In addition, for
each boxen plot, we also showed the median value of the distribution on
top for reference.

# Results

The savings estimation error ($Err$) is calculated as: $$
Err = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage ($E_{intervention}$ is the
'true' electricity usage calculated from Eq. (1) and (2)), and $E'$
indicates the estimated electricity usage either through the
conventional method or the new randomized method. Figure
\@ref(fig:abstract) shows the overall results of M&V methods comparison
and for clarity, we plotted conventional M&V results in grey, randomized
M&V results after satisfying all stopping criteria in green, and the
randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Err|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using a
time-of-week-temperature (TOWT) model [@mathieu_quantifyin

The conventional M&V method exhibits a median deviation of 4.8% in
savings estimation and 4.6% after model normalization, whereas the
randomized method demonstrates significantly smaller deviations. If the
analyst stops immediately after satisfying all stopping criteria, the
deviation is reduced to approximately 1.5% based on measured weather and
1.2% after normalization. Extending the M&V period to match the same
time period of the conventional method (24 months) further improves
accuracy to 0.7% deviation and 0.6% after normalization. Overall, we
found 83% of all filtered buildings show less error when using the
randomized M&V, and subplot (b) indicates around 74% improvement in
accuracy (median error dropped from 4.6% to 1.2%) at the time of
satisfying all stopping criteria.

Furthermore, by comparing subplots (a) and (b), we observed a minor
improvement in estimation accuracy when a model is fitted to the data to
account for weather differences. The model fitting is most beneficial
for the shorter dataset (randomized and stopping when all criteria are
met, typically 24-36 weeks of data), where it reduces deviation from
1.5% to 1.2% (around 20% improvement). This is because the weather
conditions between the first and second year for each of the locations
are relatively similar on average over this time period, and thus the
effect of adjusting for temperature as an independent variable in the
model is relatively small.

Subplot (c) compares the overall timeline of the two M&V methods. The
results show that almost all buildings achieve accurate M&V results and
meet the stopping criteria within 24 weeks (for this intervention), with
the remainder meeting all stopping criteria within 36 weeks. The
additional 12 weeks are usually driven by a need to span sufficient
weather conditions representative of a full year of data. The red arrow
highlights the excess time required by the conventional M&V method (24
months). Thus, the randomized M&V method finds more accurate results in
approximately a quarter of the time required for conventional M&V. Even
if it was the case that an existing 12-month baseline data was available
at the start of the M&V process, randomized M&V would still obtain a
more accurate result, in less time, than obtaining the 12-month retrofit
dataset required by conventional M&V.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period). Subplot (a) shows the savings estimated from measured weather conditions, and subplot b) shows the savings normalized on the typical meteorological weather of the building location after using a time-of-week-temperature (TOWT) model. Subplot (c) shows at the end of each blocking period the number of buildings satisfying all stopping criteria. The figure shows the randomized M&V method can provide much more accurate savings estimation than the convention method even at 36 weeks.", fig.height=10, fig.width=9}

#### GA ####
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p_accuracy <- ggarrange(p_top, p_middle, 
                        ncol = 1, nrow = 2,
                        labels = c("a)", "b)"),
                        common.legend = T, 
                        legend = "bottom")

# Timeline plot
df_time <- df_sprt_all_variable %>% 
  filter(seq != "final") %>% 
  bind_rows(df_sprt_all_stable %>% filter(seq != "final")) %>% 
  select(name, seq, n_weeks)

count <- list()
n <- 1
for (i in seq(0, 36, by = 3)){

  df_sprt <- df_time %>% 
    filter(seq == "sprt",
           n_weeks <= i)

  df_eob <- df_time %>% 
    filter(seq == "eob",
           n_weeks <= i)

  df_temp <- df_time %>% 
    filter(seq == "temp", 
           n_weeks <= i)

  count[[n]] <- tibble("n_weeks" = i, 
                       "eob" = nrow(df_eob), 
                       "temp" = nrow(df_temp), 
                       "sprt" = nrow(df_sprt))

  n <- n + 1
}

count <- bind_rows(count)

p_timeline <- count %>% 
  ggplot() +
  geom_bar(data = . %>% 
             filter(n_weeks == 24 | n_weeks == 36),
           aes(x = n_weeks, y = eob, fill = "Buildings finishing randomized M&V"), 
           stat = "identity", 
           position = "stack",
           alpha = 0.8, 
           width = 3) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_segment(aes(x = 36.5, y = max(eob), xend = 95.5, yend = max(eob)),
               arrow = arrow(length = unit(0.25, "in")),   
               linewidth = 1.1,  
               color = "#fb8072") + 
  annotate(geom = "text", 
           x = 66, 
           y = max(count$eob) - 35, 
           size = 5,
           label = "Excess time by conventional M&V") +
  geom_vline(xintercept = c(12, 24, 36, 48, 96), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(6, 48, by = 12), 
           y = 200, 
           size = 4, 
           label = paste0("12-week\nblock"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 100),
                     breaks = c(12, 24, 36, 48, 96), 
                     labels = c("12 weeks", "24 weeks", "36 weeks", "1 year", "2 years")) +
  coord_cartesian(ylim = c(0, 650)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       subtitle = "timeline comparison") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall savings estimation comparison by different M&V methods")

MW_summary <- df_MW_A %>% 
  pivot_wider(names_from = "method", values_from = "diff") %>% 
  group_by(name) %>% 
  summarise(improv_eob = (conv_diff - rand_eob_diff), 
            improv_final = (conv_diff - rand_final_diff)) %>% 
  ungroup() %>% 
  summarise(across(c(improv_eob, improv_final), median))

TW_summary <- df_TW_A %>% 
  pivot_wider(names_from = "method", values_from = "diff") %>% 
  group_by(name) %>% 
  summarise(improv_eob = (conv_diff - rand_eob_diff), 
            improv_final = (conv_diff - rand_final_diff), 
            improv = ifelse(improv_eob >= 0, 1, 0)) %>% 
  ungroup() %>% 
  mutate(perc = sum(improv)) %>% 
  summarise(across(c(improv_eob, improv_final, perc), median))
```

In the supplementary material, we also showed the signed/biased (instead
of absolute) estimated deviation, highlighting that the uncertainty
range associated with the conventional method is significantly larger
compared to even the randomized method that only takes 36 weeks,
according to subplot (c) in Figure \@ref(fig:abstract).

```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=7, fig.width=9, include = FALSE}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Signed error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-15, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - annual, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Signed error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(-15, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_middle, 
          ncol = 1, nrow = 2,
          labels = c("a)", "b)")) +
  plot_annotation(title = "Signed error in savings estimation by different M&V methods")

percentiles <- df_TW_A %>%
  filter(method == "conv_diff") %>% 
  summarise(
    Q1     = round(quantile(diff, 0.25, na.rm = TRUE), digits = 2),
    Median = round(quantile(diff, 0.50, na.rm = TRUE), digits = 2),
    Q3     = round(quantile(diff, 0.75, na.rm = TRUE), digits = 2)
  )
```

Granderson et al. (2016), using over 500+ commercial buildings (on a
different open-source dataset), compared the predictive performance of a
variety of M&V baseline models from simple weekly mean to more complex
machine learning methods using the normalized mean bias error (NMBE)
[@granderson_accuracy_2016]. Since we also calculated the normalized
savings estimation error, we compared our normalized accuracy for 600+
buildings with the Time-Of-Week and Temperature (TOWT) model and Typical
Meteorological Year (TMY) weather against their TOWT model prediction
assessment in Table 2. The model and overall building datasets are
performing similarly.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|-----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

: Table 2. Savings accuracy comparison with a similar literature

## Non-routine events impact

As mentioned in Section \@ref(filter-and-clean-dataset), he majority of
the selected buildings are affected by operational drift, which is the
reason why we observe baseline electricity usage changes throughout the
two-year period. To evaluate the robustness of the two M&V methods to
the effect of operational drift and non-routine events, we repeated the
M&V run on the original dataset prior to the application of the reset
control. In other words, there is no intervention applied here, so the
most reliable M&V method will be the one that detects the closest to 0
kW savings using the measurements. Therefore, the $Err$ here are
calculated as:

$$
Err = 0-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$

Figure \@ref(fig:nosaving) shows that the randomized M&V method (both at
the time of satisfying all stopping criteria and at the end of a
two-year period) produced consistent results. We included the absolute
error ($|Err|$) figure in the supplementary material. Similarly to
Figure \@ref(fig:abstract), stopping when meeting all stopping criteria
generally yields slightly larger uncertainty than continuing for 24
months. However, it still yields a more accurate result than the
conventional method. Interestingly, similar to
[@granderson_accuracy_2016] where they obtained a median normalized
deviation around -1.25%, our analysis on the conventional method shows
that the typical deviation is -1.5%, meaning that this method would
predict that, in the absence of any known intervention, the typical
building is decreasing energy use slightly in this dataset. Meanwhile,
through random sampling, the influence of non-routine events is much
closer to 0 with narrower uncertainty range.

```{r nosaving, fig.cap="Comparison between the two M&V methods in detecting no intervention effect when buildings are subject to static non-rountine events and gradual operational drift.", fig.height=4, fig.width=9}

#### MEAN-NULL ####
rand_eob_S <- df_seq_fs_null_stable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_null_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_null_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_seq_fs_null_variable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = 0 - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_null_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - conv, 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_null_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = true - rand, 
         method = "rand_final") %>% 
  select(name, method, diff)

df_MW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

# plot for combined dataset
rand_eob_A <- bind_rows(rand_eob_V, rand_eob_S) %>% 
  select(name, rand_eob_diff = diff)

conv_A <- bind_rows(conv_V, conv_S) %>% 
  select(name, conv_diff = diff)

rand_final_A <- bind_rows(rand_final_V, rand_final_S) %>% 
  select(name, rand_final_diff = diff)

df_MW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                                "conv_diff" = "Conventional", 
                                "rand_eob_diff" = "Randomized", 
                                "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 6), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Estimated effect in fractional savings",
       title = "Estimated effect of non-routine events by different M&V methods", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(-16, 16)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

# Sensitivity analysis

## TOWT modeling accuracy

To evaluate modeling accuracy, we used the Coefficient of Variation of
Root-Mean Squared Error (CV(RMSE)) as the primary error metric. CV(RMSE)
is a normalized measure, thus it enables direct comparison across
different modeling results. Since we have the ground-truth baseline
measurements in the post-retrofit period, we can assess the error caused
by model adjustment (i.e., how accurate is the 'counterfactual' baseline
provided by the two M&V methods). Figure \@ref(fig:towtacc) shows the
difference between the true post-retrofit baseline and adjusted baseline
through TOWT modeling; since the quantity of the training set is the
same, this highlights the impact of the data sampling technique. For
example, if a non-routine event happens (such as a tenant occupying an
entire floor of a building moving out), the TOWT model can yield
significant deviations. This is particularly the case for the
conventional M&V method since it samples continuously throughout the
pre-retrofit period. However, randomized sampling is less impacted by
these events since it samples only 50% of the pre-retrofit baseline
throughout a blocking period. Consequently, the figure shows there is a
noticeable improvement in modeling accuracy when using the randomized
method.

```{r towtacc, fig.cap="TOWT modeling accuracy distribution for all buildings included (with each data point representing one building)", fig.width=9, fig.height=4}

cvrmse <- df_model_acc_stable %>% 
  bind_rows(df_model_acc_variable) %>% 
  pivot_longer(c(conv, rand), 
               names_to = "method", 
               values_to = "cvrmse") %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, 
                              "rand" = "Randomized\n(24 months)", 
                              "conv" = "Conventional"))

result <- t.test(filter(cvrmse, method == "Randomized\n(24 months)")$cvrmse, 
                 filter(cvrmse, method == "Conventional")$cvrmse, 
                 exact = TRUE,
                 alternative = "less",
                 mu = 0,
                 conf.int = TRUE, 
                 conf.level = 0.90,
                 var.equal = TRUE)$p.value

cvrmse %>% 
  ggplot(aes(x = method, y = cvrmse, fill = method)) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method), linewidth = 0.25) +
  annotate(geom = "text", 
           x = 1.5, 
           y = 40, 
           size = 4, 
           label = ifelse(result < 0.001, "P value < 0.001", paste0("P value = ", result))) +
  geom_errorbarh(aes(xmin = 1.2, xmax = 1.8, y = 38), height = 2, linewidth = 0.25) +
  geom_text(data = . %>% 
              group_by(method) %>% 
              summarise(median = median(cvrmse, na.rm = T)) %>% 
              ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 0), " %")), 
            size = 5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %")) +
  coord_cartesian(ylim = c(0, 42)) +
  labs(x = NULL, 
       y = "CV(RMSE)", 
       color = NULL, 
       title = "TOWT model fitting accuracy by different M&V methods") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Sampling interval impact

In some cases, particularly those involving large thermal mass and
active thermal energy storage systems or operating continuously without
shutting down for switchover, there can be a substantial time-lagged
effect associated with interventions. This means that the effect of one
control strategy may persist, or 'carryover', after switching to
another. To mitigate such carryover effects, building analysts may
choose to drop a predetermined number of non-consecutive days in the
switchback experiment. For example, consider a heat pump water heater
where the tank is charged at very different times of the day by the
intervention strategy compared to the baseline strategy. When switching
the strategy, the tank may be much warmer or cooler than typical for
that strategy, thus influencing subsequent measurements, particularly
the performance measured on the first day (i.e., a non-consecutive day).
In such cases, increasing the sampling interval (e.g., sample strategies
every 2 or 3 days instead of daily) and discarding measurements from
those non-consecutive days can eliminate carryover. However, this
approach reduces the total data collected: for instance, sampling every
3 days and dropping days with non-consecutive control strategies would
drop 1/6 of measurements on average. Sampling every 2 days means
dropping a quarter of the measurements. Therefore, by increasing the
sampling interval, there is an accuracy penalty due to a reduction in
randomization (how often samples occur); when decreasing the sampling
interval, there is an accuracy penalty from dropping more
non-consecutive days. The dataset contains no specific information about
the buildings' thermal mass condition. So, to quantify such trade-off,
we compared the accuracy of different sampling intervals and the
consequence of dropping the non-consecutive days at the time when
randomized M&V produces a result (after satisfying all stopping
criteria). These results are shown in Figure \@ref(fig:absinterval).

```{r absinterval, fig.cap="Comparison of different sampling interval impact on M&V estimation accuracy (dropped: all non-consecutive days were dropped; kept: all measurements were kept)", fig.height=8, fig.width=9}

#### ABS-INT ####
# sprt sequence plot for different sampling intervals
df_seq_interval_fs_keep_stable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_seq_fs_stable %>% 
              filter(seq == "eob") %>% 
              select(-seq) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "fs"))

df_seq_interval_fs_keep_variable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_seq_fs_variable %>% 
              filter(seq == "eob") %>% 
              select(-seq) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "fs"))

df_drop <- df_seq_interval_fs_drop_stable %>% 
  bind_rows(df_seq_interval_fs_drop_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "drop")

df_keep <- df_seq_interval_fs_keep_stable %>% 
  bind_rows(df_seq_interval_fs_keep_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "keep")

df_MW_A <- df_drop %>% 
  bind_rows(df_keep) %>% 
  left_join(rbind(df_fs_stable %>% 
                    filter(scenario == "ref" & method == "true"), 
                  df_fs_variable %>% 
                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(diff = abs(savings - eob)) %>% 
  filter(is.finite(diff))

p_top <- df_MW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, inf.rm = T)) %>% ungroup(), 
            aes(x = interval, y = median, group = group,
                label = paste0(round(median, digits = 1), " %")), 
            position = position_dodge(width = 0.75), 
            color = "white", 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 12)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# with tmy version
df_seq_interval_nm_keep_stable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_sprt_all_stable %>% 
              filter(seq == "eob") %>% 
              select(-c(seq, n_weeks)) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "annual"))

df_seq_interval_nm_keep_variable %<>% 
  select(-c(sprt, temp, final)) %>% 
  bind_rows(df_sprt_all_variable %>% 
              filter(seq == "eob") %>% 
              select(-c(seq, n_weeks)) %>% 
              mutate(interval = 1) %>% 
              rename(eob = "annual"))

df_drop <- df_seq_interval_nm_drop_stable %>% 
  bind_rows(df_seq_interval_nm_drop_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "drop")

df_keep <- df_seq_interval_nm_keep_stable %>% 
  bind_rows(df_seq_interval_nm_keep_variable) %>% 
  select(name, eob, interval) %>% 
  mutate(group = "keep")

df_TW_A <- df_drop %>% 
  bind_rows(df_keep) %>% 
  left_join(rbind(df_fs_stable %>% 
                    filter(scenario == "ref" & method == "true"), 
                  df_fs_variable %>% 
                    filter(scenario == "ref" & method == "true")), by = "name") %>% 
  mutate(diff = abs(savings - eob)) %>% 
  filter(is.finite(diff))

p_bottom <- df_TW_A %>% 
  mutate(interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept")) %>% 
  ggplot(aes(x = interval, y = diff, fill = group)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = group), width = 0.8, linewidth = 0.15) +
  geom_text(data = . %>% group_by(group, interval) %>% summarise(median = median(diff, na.rm = T)) %>% ungroup(), 
            aes(x = interval, y = median, group = group,
                label = paste0(round(median, digits = 1), " %")), 
            position = position_dodge(width = 0.75), 
            color = "white", 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 5), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 12)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p_top, p_bottom, 
          ncol = 1, nrow = 2, 
          align = "hv", 
          labels = c("a)", "b)"),
          common.legend = T, 
          legend = "bottom") +
  plot_annotation(title = "Absolute error in savings estimation of randomized M&V\nafter satisfying all stopping criteria")
```

To quantify the impact of reduced randomness, all measurements were kept
when calculating the distribution shown in the darker-colored set. The
results suggest that without dropping non-consecutive days in subplot
(a), there was a gradual increase in error from 1.5% to 2.3% (1.2% to
1.9% with weather normalization in subplot (b)) when the sampling
interval increases from daily to weekly intervals. This is due to the
fact that increasing the sample interval means reducing randomization.

To quantify the combined impact of data loss due to preventing the
carryover effect, we generated the lighter-colored set, which drops all
non-consecutive days. As a result, we noticed an increase in estimation
error when sampling daily to 2.4% (to 1.9% when TOWT is fitted), which
led to a similar accuracy compared to weekly intervals. Therefore,
considering this trade-off, we recommend using a two-day sampling
interval (or three-day sampling interval if the building owner prefers
less frequent switchover) if the carryover effect is likely but expected
to last less than one day, and then normalizing the estimation via
energy modeling.

The supplementary material contains similar figures showing biased
deviation distribution, as well as the same results when continuing the
randomized M&V for 24 months. Those results indicate the same
conclusion; that sampling every two days and dropping non-consecutive
samples yields the optimal result when carryover effects are present.

Sampling at different intervals and excluding non-consecutive days also
impact other stopping criteria, such as the time needed to satisfy
statistical uncertainty and weather conditions thresholds. In Figure
\@ref(fig:intervaltime), we compared the percentage of buildings
satisfying all stopping criteria under different sampling intervals,
with or without dropping non-consecutive days. The results indicate that
randomized M&V concludes more quickly when all data are retained,
particularly at shorter intervals, as expected. For instance, when
sampling at a one-day interval, 90% of buildings meet all criteria
within 24 weeks if no data are dropped, but this number decreases
significantly to 52% when non-consecutive days (account for 1/3 of the
data) are excluded, and 3% of the buildings require another 12 weeks to
satisfy the weather criterion. As the sampling interval increases, this
difference gradually diminishes: when sampling at a weekly interval,
removing non-consecutive days has no impact on the timeline of meeting
stopping criteria.

```{r intervaltime, fig.cap="Comparison of time required to satisfy all stopping criteria when considering dropping non-consecutive days (or not)", fig.height=7, fig.width=9}
df_drop <- bind_rows(df_seq_interval_tl_drop_stable, df_seq_interval_tl_drop_variable) %>%
  mutate(group = "drop", 
         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
  select(name, seq, n_weeks, interval, group)

interval_1_keep <- bind_rows(df_sprt_all_stable, df_sprt_all_variable) %>%
  select(c(name, seq, n_weeks)) %>%
  filter(seq != "final") %>%
  mutate(seq = as.factor(seq)) %>%
  mutate(interval = 1, 
         group = "keep")

df_keep <- bind_rows(df_seq_interval_tl_keep_stable, df_seq_interval_tl_keep_variable) %>%
  mutate(group = "keep", 
         temp = ifelse(base_temp >= interv_temp, base_temp, interv_temp)) %>% 
  pivot_longer(c(sprt, temp, eob, final), names_to = "seq", values_to = "n_weeks") %>% 
  select(name, seq, n_weeks, interval, group) %>% 
  bind_rows(interval_1_keep)

df_MW_A <- bind_rows(df_keep, df_drop) %>% 
  filter(seq != "final") %>% 
  filter(seq == "eob") %>% 
  mutate(seq = as.factor(seq), 
         seq = recode_factor(seq, 
                             "eob" = "Buildings finishing all criteria"), 
         interval = as.factor(interval), 
         interval = recode_factor(interval, 
                                  "1" = "1-day\nsampling",
                                  "2" = "2-day\nsampling", 
                                  "3" = "3-day\nsampling", 
                                  "7" = "7-day\nsampling"), 
         group = as.factor(group), 
         group = recode_factor(group, "drop" = "Dropped", "keep" = "Kept"), 
         n_weeks = as.factor(n_weeks), 
         n_weeks = recode_factor(n_weeks, 
                                 "12" = "12\nweeks", 
                                 "24" = "24\nweeks", 
                                 "36" = "36\nweeks", 
                                 "48" = "48\nweeks")) %>% 
  group_by(n_weeks, group, interval) %>% 
  summarise(n = n()) %>%
  ungroup()

levA <- levels(df_MW_A$n_weeks)
levB <- levels(df_MW_A$group)
levC <- levels(df_MW_A$interval)

# 2. Expand to all combinations of factorA and factorB
all_combinations <- expand_grid(
  n_weeks = levA,
  group = levB, 
  interval = levC, 
)

df_MW_A <- all_combinations %>%
  left_join(df_MW_A, by = c("n_weeks", "group", "interval")) %>% 
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  group_by(group, interval) %>% 
  mutate(n = ifelse(n > sum(n) * 0.01, n, 0), 
         perc = n / sum(n) * 100) %>% 
  ungroup()

df_MW_A %>% 
  ggplot(aes(x = n_weeks, y = perc, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~interval, nrow = 2) +
  geom_text(aes(x = n_weeks, y = perc + 5, label = paste0(round(perc, digits = 0), "%"), group = group), 
            position = position_dodge(1), 
            size = 5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Percentage of buildings satisfying all stopping critiera", 
       subtitle = "Counted at the end of each 12-week blocking period") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

## Sampling ratio impact

The sampling ratio determines the likelihood that the M&V analyst will
apply the baseline or intervention, and we recommend using an equal
sampling ratio at the outset of the M&V process. In our previous study
[@raftery2024], we also recommended that building owners adjust the
sampling ratio by reducing the amount of baseline sampling after meeting
all stopping criteria. This is because, although sampling the baseline
allows for tracking changes within buildings, it can also limit
opportunities for building owners to implement the intervention, thereby
reducing potential savings. To further examine its impact on M&V
accuracy, Figure \@ref(fig:abscont) compares two sampling ratios
(baseline/intervention: 20%/80% and 10%/90%) and two associated
scenarios with the previously shown results when sampling at 50%/50%.
The "Continued" group (shown in a lighter color) follows the standard
randomized M&V framework with an initial 50%/50% sampling ratio, then
switches to the indicated sampling ratio after satisfying the stopping
criteria, continuing this until the end of the two-year period. In
contrast, the "Randomized" group uses the indicated ratio from the
outset and continues throughout the two years. The figure below shows
that absolute error is significantly lower without the initial 50%/50%
sampling, due to baseline drift and non-routine events that most
buildings experience over two years (as noted in Section
\@ref(filter-and-clean-dataset)). This is also the case even with
weather normalization, the results of which are included in the
supplementary material.

```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=4, fig.width=9}

#### ABS-CONT ####
df_MW_S <- df_cont_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs), 
         diff_all = abs(savings - cont_fs_all)) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_stable %>% 
              filter(method != "conv") %>% 
              pivot_wider(names_from = method, values_from = savings) %>% 
              mutate(values = abs(true - rand), 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for the variable subset
df_MW_V <- df_cont_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - cont_fs), 
         diff_all = abs(savings - cont_fs_all)) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_variable %>% 
              filter(method != "conv") %>% 
              pivot_wider(names_from = method, values_from = savings) %>% 
              mutate(values = abs(true - rand), 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for combined dataset
df_MW_A <- rbind(df_MW_S, df_MW_V)

df_MW_A %>% 
  mutate(ratio = as.factor(ratio), 
         ratio = recode_factor(ratio, 
                               "0" = "Randomized\n(50/50)", 
                               "1_1" = "Continued\n(20/80)", 
                               "1_2" = "Randomized\n(20/80)", 
                               "2_1" = "Continued\n(10/90)", 
                               "2_2" = "Randomized\n(10/90)")) %>% 
  ggplot(aes(x = ratio, y = values, fill = ratio)) +
  geom_lv(k = 4, outlier.shape = NA, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.1)) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = ratio), width = 0.75, linewidth = 0.15) +
  geom_vline(xintercept = 1.5, color = "#fb8072", linewidth = 1, lty = "dashed") +
  geom_text(data = . %>% 
              group_by(ratio) %>% 
              summarise(median = median(values)) %>% 
              ungroup(), 
            aes(x = ratio, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 5) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       title = "Absolute error in savings estimation by different sampling ratio",
       subtitle = str_glue("All {A_building} buildings with measured weather conditions")) +
  coord_cartesian(ylim = c(0, 15)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

We further illustrated this point in the supplementary material where we
plotted results using only the 'stable' building subset (66 buildings),
as defined in Section \@ref(filter-and-clean-dataset). Since they are
less influenced by non-routine events, the differences between sampling
strategies are much reduced in this subset and can be largely corrected
through weather normalization modeling.

As shown in the previous overall M&V method comparison Figure
\@ref(fig:abstract) in the Results section (plotted again on the left
side of the red dashed line), the highest accuracy is achieved when
sampling at a 50%/50% ratio. Therefore, we continue to recommend that
building owners follow the proposed framework and flow chart from our
previous study [@raftery2024], and maintain a 50%/50% sampling ratio
until all stopping criteria are met. If building owners choose to adjust
the sampling ratio afterwards and a change in the building baseline is
detected, the updated savings estimation should exclude the baseline
measured prior to the sampling ratio change. As shown in the figure,
additional baseline measurements before the ratio change do not
contribute to reducing errors or uncertainty. The results also indicate
that, if the building owner plans to conduct a two-year test and aims to
maximize savings, it is recommended to implement a 10%/90% sampling
ratio between the baseline and intervention from the outset.

## Blocking period impact

A blocking period refers to the given period during which an equal
number of baseline and intervention days are sampled, and the associated
day-of-week factors are balanced between the two. Selecting different
blocking periods affects how frequently the M&V analyst evaluates
stopping criteria and may influence the decision when the saving is
detected at a sufficient level of certainty. Here, we summarize the
advantages and disadvantages of shorter versus longer blocking periods.
The second step of the randomized M&V method involves choosing an
appropriate blocking period to minimize potential biases caused by
incomplete randomization. For most practical scenarios, we recommend a
blocking period of approximately 12 to 16 weeks, as it typically covers
the duration of a season, ensuring a representative sampling of
operational conditions. In general, longer blocking periods are
advantageous when evaluating multiple intervention strategies or using
sampling intervals greater than one day. In such cases, extended
blocking ensures a balanced distribution of both day-of-week and outdoor
weather conditions.

Opting for a shorter blocking period, such as 6 weeks, has the potential
benefit of allowing the stopping criteria to be met earlier. We assessed
the use of a 6-week blocking period to see whether buildings finished
earlier in the original 12-week blocking period can reduce the time
required with a shorter blocking period. We plotted the percentage of
buildings finishing all stopping criteria when the blocking period is
set as 6 weeks in Figure \@ref(fig:6weekblock). When compared to Figure
\@ref(fig:intervaltime), the only difference we found is when sampling
at a daily interval and not dropping non-consecutive days, 12% of
buildings can get an estimate earlier. This means those buildings are
likely to meet all stopping criteria between 12 weeks and 18 weeks but
when sampled over a 12-week block, the savings can only be determined at
week 24. Although sampled baseline and intervention days can be
different when the blocking period is changed, the results indicate most
buildings still require 9 months to meet outdoor weather conditions when
starting M&V from January.

```{r 6weekblock, fig.cap = "Time required to satisfy all stopping criteria when using a 6-week blocking period"}

knitr::include_graphics(paste0(fig_path, "6week_time.png"))
```

Sampling at a shorter blocking period also means that the analyst must
perform the hypothesis tests more frequently, which can necessitate
adjusting the significance level using methods such as the Bonferroni
correction to control for increased type-I error risk. On the plus side,
a shorter blocking period can reduce the adverse impact of missing or
problematic data. For example, if an intervention requires fine-tuning
after initial implementation, removing data from the first blocking
period would be less detrimental compared to longer blocking periods,
preserving the overall integrity of the experiment. Based on these
analyses we recommend to use a blocking period of 6 weeks when sampled
at a daily interval, but a longer blocking period such as 12 weeks when
a longer sampling interval is required.

## M&V starting time impact

In this study, we set the baseline period to begin on January 1st, which
offers the advantage of capturing colder outdoor weather conditions
early in the M&V process. However, when the baseline measurements miss
the initial winter period, meeting the required weather condition
criteria can potentially take a longer time. A previous study also
showed that M&V accuracy, especially model fitting, depends on both
monitoring duration and starting month, with optimal modeling achieved
when sufficient weather variation is captured [@singh_predicting_2014].
To assess this impact in our switchback randomization, we repeated the
randomized M&V analysis for all buildings using a new start date in the
swing season in March. As shown in Figure \@ref(fig:startT), the new M&V
process results in a noticeable delay in satisfying the stopping
criteria. Compared to Figure \@ref(fig:intervaltime), where a January
start allowed most buildings to finish within 36 weeks (\~8 months), a
March start extended the required duration to 48 weeks (\~11 months) for
the majority of buildings. Furthermore, when a longer sampling interval
is used or non-consecutive days are excluded (to avoid carryover
effects), some buildings may require up to 60 weeks (\~14 months) to
complete the randomized M&V. As expected, to collect sufficient data
representing winter conditions, the sampling may need to continue until
the next January (around week 48). However, as the overall data
collection period extends, the influence of dropping non-consecutive
days diminishes and becomes less of a concern.

```{r startT, fig.cap = "Time required to satisfy all stopping criteria starting from March"}

knitr::include_graphics(paste0(fig_path, "startT.png"))
```

Similarly, we evaluated the time required to satisfy all stopping
criteria when using a shorter blocking period, and included the plot in
the supplementary material. We found that compared to Figure
\@ref(fig:startT), a significant proportion of buildings completed the
randomized M&V by week 42 (mid-December), which is one blocking period
earlier than having a 12-week blocking period. This result further
highlights that, in general, starting M&V in the middle of the winter or
summer periods will yield results spanning the full range of weather
conditions faster than starting at other times of year, such as during a
swing season.

## Limitations

We identify two key limitations in this study. First, the simulated
intervention (chilled water temperature reset) remains somewhat generic
and simplistic, considering the diverse building types and climate zones
in the BDG2 dataset. As the primary focus is to accurately detect the
intervention (or any intervention) effect so for simplicity, we applied
the parameters listed in Table 1 uniformly across all buildings.
Secondly, for the main analysis, we assumed no carryover effect and that
a daily sampling interval would suffice for most commercial buildings,
but certainly exceptions exist. However, in the sensitivity analysis, we
addressed such concern by comparing different sampling strategies and
the impact of dropping measurements from non-consecutive days.

# Conclusion

We assessed the performance of a randomized whole-building measurement
and verification (M&V) method by comparing it to the conventional M&V
approach described in the IPMVP and in ASHRAE Guideline 14, using a
large, open-source commercial building dataset. The randomized M&V
method leverages the randomized experimental design concept from other
scientific fields, along with statistical sequential inference
techniques, to determine when savings are detected at sufficient
certainty. We used a virtual control retrofit case, resetting the
chilled water setpoint based on outdoor temperature, and applied it to
over 600 commercial buildings. By comparing the savings estimations of
the conventional method with the novel randomized method, we found that
the randomized approach provides faster and more robust savings
estimations.

Specifically, we showed that throughout 11 different locations assessed
in this study, the randomized M&V can provide a savings estimation after
36 weeks (with the majority finished by 24 weeks) once all stopping
criteria are satisfied. In contrast, the conventional method requires a
full range of baseline and intervention measurements under normal
operating conditions, typically requiring each operational strategy to
be measured for a full year. Moreover, we verified that, even if the
randomized M&V requires less time, it is more accurate showing an
absolute error of 1 - 2%. Whereas for the much longer two-year
conventional method, the estimation error was approximately 5%.

We also evaluated the impact of non-routine events, particularly more
gradual and subtle operational drift, on the proposed M&V method; the
results show that baseline changes in the post-retrofit period are very
common for realistic buildings and can deviate savings estimated using
the conventional method. In contrast, we found that those events have a
very negligible impact on the savings estimated using the randomized
method, demonstrating robustness to this issue.

Furthermore, we considered scenarios where there is a known carryover
effect from switching between strategies in the building and assessed
the impact of dropping samples when the strategy is non-consecutive and
varying the sampling interval. Through that process, we showed that for
the typical building, when carryover is present, using a 2-day sampling
interval and excluding days with non-consecutive control strategies
operating yields the optimal design for the median building in this
dataset.

Lastly, we investigated the sensitivity of time required to complete the
randomized M&V method under different scenarios. We found that measuring
a sufficient range of outdoor weather conditions remains the most
stringent criterion. However, even in the most non-ideal case, the
randomized method can still produce an estimate much faster than two
years and  and with significantly lower error.

# CRediT authorship contribution statement

**Aoyu Zou**: Conceptualization, Data curation, Formal analysis,
Methodology, Investigation, Software, Writing - original draft,
Writing - review & editing. **Paul Raftery**: Conceptualization, Funding
acquisition, Formal analysis, Methodology, Investigation, Project
administration, Supervision, Writing - review & editing. **Stefano
Schiavon**: Conceptualization, Methodology, Visualizations,
Investigation, Supervision, Writing - review & editing. **Carlos
Duarte**: Methodology, Investigation, Supervision, Writing - review &
editing. **Gail Brager**: Supervision, Writing - review & editing.

# Reproducibility

A reproducible example with analysis code is available (MIT license) at
<https://github.com/CenterForTheBuiltEnvironment/genome_mnv>.

# Declaration of competing interest

The Center for the Built Environment (CBE) at the University of
California, Berkeley, with which the authors are affiliated, is advised,
and funded in part, by many partners that represent a diversity of
organizations from the building industry, including manufacturers,
building owners, facility managers, contractors, architects, engineers,
government agencies, and utilities.

# Acknowledgements

# References